### 2017-10-18 paper/R/webhoseio.Rmd

---
title: "Topic Modeling of News"
author: "Franziska Löw"
date: "`r format(Sys.Date())`"
output: github_document
---

```{r, include=FALSE}
# load the packages
libs <- c("tidytext","tidyr","readr","lubridate","tm","stm","RColorBrewer",
          "plyr","dplyr","class","knitr","kableExtra","cldr","data.table",
          "htmlTable","ggplot2","gridExtra","jsonlite","stringr","scales","rjson")
lapply(libs, library, character.only = TRUE)

```

# Load and prepare Dataframe

```{r, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls())

# create a character vector, with each file name represented by an entry
#june <- list.files("~/CloudStation/textmining/data/webhoseio/2017-06", pattern="*.json", full.names=TRUE) 
#july <- list.files("~/CloudStation/textmining/data/webhoseio/2017-07", pattern="*.json", full.names=TRUE) 
aug <- list.files("~/CloudStation/textmining/data/webhoseio/2017-08", pattern="*.json", full.names=TRUE) 
#sep <- list.files("~/CloudStation/textmining/data/webhoseio/2017-09", pattern="*.json", full.names=TRUE)

# a list in which each element is one of your original JSON files
#myJSON <- lapply(c(june,july,aug,sep), function(x) fromJSON(file=x)) 
myJSON <- lapply(aug, function(x) fromJSON(file=x))

length(myJSON)
```

## Convert List to Dataframe

```{r, eval=FALSE, include=FALSE}
site <- NULL
title <- NULL
text <- NULL
published <- NULL
fb_shares <- NULL
fb_likes <- NULL
section <- NULL
external_links <- NULL

for (i in 1:length(myJSON)){
  site <- append(site, myJSON[[i]]["thread"][[1]]["site"][[1]])
  title <- append(title, myJSON[[i]]["title"][[1]])
  text <- append(text, myJSON[[i]]["text"][[1]])
  published <- append(published, myJSON[[i]]["published"][[1]])

  fb_shares <- append(fb_shares, myJSON[[i]]["thread"][[1]]["social"][[1]]["facebook"][[1]][["shares"]])
  fb_likes <- append(fb_likes, myJSON[[i]]["thread"][[1]]["social"][[1]]["facebook"][[1]][["likes"]])

  section <- append(section, myJSON[[i]]["thread"][[1]]["section_title"][[1]])
  external_links <- append(external_links, length(myJSON[[i]]["external_links"][[1]]))

}

df <- cbind(site, title, text, published, fb_shares, fb_likes, section, external_links)
df <- data.frame(df, stringsAsFactors = FALSE)

rm(site, title, text, published, fb_shares, fb_likes, section, external_links)
```

## Define private and public service operators

```{r, eval=FALSE, include=FALSE}
private <- c("bild.de","focus.de","handelsblatt.com","spiegel.de","stern.de","welt.de","zeit.de")

df %>%
  mutate(owned = ifelse(site %in% private, "private", "public")) -> df
```

## Subset by
```{r, eval=FALSE, include=FALSE}
# .... date
df$date <- substr(df$published, 1,10)
df$date <- as.POSIXct(df$date, format ="%Y-%m-%d")

df <- subset(df, df$date > "2017-09-01")
```

```{r, eval=FALSE, include=FALSE}
# ... content
content <- c("Bundestagswahl","Merkel","Schulz","Wagenknecht","Lindner","Gauland","Özdemir",
             "CDU","SPD","FDP","AfD","Die Grünen","Die Linke")

btw <- df[which(grepl(paste(content, collapse = "|"), df$title)),]

# Save file
write.csv(btw, file="~/CloudStation/textmining/R/output/btw.csv")
```

## Load Data

```{r}
rm(list = ls())
df <- read.csv("output/btw.csv")
```

### Number of published Articels by News Source

```{r}
col <- brewer.pal(12, name = "Paired")

df %>%
  group_by(site) %>%
  tally(sort = TRUE) %>%
  ggplot(aes(site, n)) +
  geom_col(fill = col, alpha = .8) +
  xlab("") +
  ylab("") +
  theme(axis.text.x = element_text(angle = 40, hjust = 1))
```

### Number of published Articels by Ownership

```{r}
df %>%
  group_by(owned) %>%
  tally(sort = TRUE) %>%
  ggplot(aes(owned, n)) +
  geom_col(fill = col[c(6,2)], alpha = .8) +
  geom_text(aes(label=n), position = position_stack(vjust = 0.5)) +
  xlab("") +
  ylab("") +
  theme(axis.text.x = element_text(angle = 40, hjust = 1))
```

# Pre-Process Text

```{r, echo=FALSE, message=FALSE, warning=FALSE}
clean.text = function(x)
  {
  x = gsub("[[:punct:]]", " ", x)  # replace punctuation with space

  x = gsub("DIE WELT", "", x)
  x = gsub("SPIEGEL ONLINE", "", x)
  x = gsub("FOCUS Online", "", x)
  x = gsub("ZEIT ONLINE", "", x)
  x = gsub("stern de", "", x)
  x = gsub("Handelsblatt", "", x)
  x = gsub("Politik Inland Bild de", "", x)
  x = gsub("BILDplus Inhalt", "", x)
  x = gsub("Bild de", "", x)
  x = gsub("SWR Aktuell", "", x)
  x = gsub("SWR de", "", x)
  x = gsub("[ ]NDR[ ]", "", x)
  x = gsub('ZDF \"\"Morgenmagazin\"\"', "", x)
  x = gsub("ZDFmediathek", "", x)
  x = gsub("[ ]ZDF[ ]", "", x)
  x = gsub("[ ]ARD[ ]", "", x)

  x = gsub("die neuesten Nachrichten", "", x)
  x = gsub("RSS Feed", "", x)
  x = gsub("Der Überblick   7 mal 17   Alle Artikel", "", x)
  x = gsub("Der Tag im Überblick", "", x)
  x = gsub("Inhalt Seite 1", "", x)
  x = gsub("Seite 2", "", x)
  x = gsub("dpa", "", x)
  x = gsub("Twitter Anzeige", "", x)
  x = gsub("Anzeige", "", x)
  x = gsub("Foto  \\w+", "", x)
  x = gsub("Video", "", x)
  x = gsub("Quelle \\w+", "", x)
  x = gsub("\\w+ \\w+ Getty Images", "", x)
  x = gsub("AFP", "", x)
  x = gsub("August", "", x)
  x = gsub("Live-Ticker", "", x)
  x = gsub("Deutsche Presse Agentur GmbH   Nachrichtenagentur Alle Beiträge von  anzeigen", "", x)
  x = gsub("Deutsche Presse Agentur", "", x)

  x = gsub("[[:punct:]]", " ", x)  # replace punctuation with space
  x = gsub("[[:cntrl:]]", " ", x)  # replace control characters with space
  x = gsub("[[:digit:]]", "", x)  # remove numbers
  x = gsub("^[[:space:]]+", "", x) # remove whitespace at beginning of documents
  x = gsub("[[:space:]]+$", "", x) # remove whitespace at end of documents 
  x = gsub("[\n{2,}]", "", x)
  x = tolower(x)
  return(x)
}

# apply function to the "orig" dataframe
df$text_cleaned <- clean.text(df$text)

df %>%
  filter(grepl("Getty Images", text)) %>%
  select(text, text_cleaned) %>%
  htmlTable(align = "l")
```

```{r}
# get stopwords from online dict and add costum stopwords
#exceptions   <- c("nicht")
#my_stopwords <- setdiff(stopwords("german"), exceptions)
my_stopwords <- c(stopwords("german"), "spon","zeitmagazin","sz", "seitennavigation","startseite","gibt","immer","uhr","dass","mehr","sagte","sei","wurde","twitterfacebookgoogle","whatsappteilen","kartengeschichte","elbvertiefung","econa","prozent","freepik","colourbox")

# Remove stopwords
df$text_cleaned<- removeWords(df$text_cleaned, my_stopwords)

# save file
write.csv(df, file = "/Users/Franzi/CloudStation/textmining/data/news_cleaned.csv")
```

## Tokenize terms
```{r}
news.token <- df %>%
  group_by(site) %>%
  unnest_tokens(word, text_cleaned) %>%
  count(site, word, sort = TRUE)  %>%
  bind_tf_idf(word, site, n) %>%
  arrange(desc(tf_idf)) 
```

# Topic Modeling
## LDA
```{r}
# Build corpus
btw.lda <- Corpus(VectorSource(df$text_cleaned))
# Create document-term matrix
dtm <- DocumentTermMatrix(btw.lda)
```

```{r}
# Run LDA using Gibbs sampling
library(topicmodels)

k = 20 # Topics

t1 <- Sys.time()
lda20 <-LDA(dtm, 20, method="Gibbs", control=list(nstart=5, iter = 10000))
t2 <- Sys.time()
t2 - t1
```

## Structural Topic Model

```{r, eval=FALSE, include=FALSE}
# Process data
processed <- textProcessor(df$text_cleaned, metadata = df[,c("date","site","text_cleaned","text")],
                           wordLengths = c(2,Inf),
                           lowercase = F,
                           removestopwords = F,
                           removenumbers = F,
                           removepunctuation = F,
                           stem = F)
out <- prepDocuments(processed$documents, processed$vocab, processed$meta)
```

### Run the Model 
```{r}

t1 <- Sys.time()
stm <- stm(documents = out$documents, 
           vocab = out$vocab, 
           K=25, 
           prevalence = ~site, 
           content = ~site, 
           data=out$meta, 
           init.type = "Spectral")
t2 <- Sys.time()
t2 - t1

save(stm, out, file = "stm.RDa")
#load(file = "stm.RDa")

```

### Interpret the results

#### Wordclouds by topic
```{r}
cloud(stm, topic =8, scale = c(5,.25))
```

```{r}
# get document - topic distribution 
theta <- as.data.frame(stm$theta)
theta$source <- out$meta$source
theta$date <- out$meta$date
```

#### get the mean distribution of topics  ...
##### ... by date
```{r}
theta %>% 
  select(- source) %>%
  group_by(date) %>% 
  summarise_all(funs(mean)) -> date_mean

# plot 
melt1 = melt(date_mean, id = "date")

ggplot(melt1, aes(x=date, y = variable)) +   
  geom_raster(aes(fill=value)) + 
  scale_fill_gradient(low = "light blue", high = "dark blue") +
  labs(x="Date", y="Topics", title="Average Topic Weights") + 
  theme_bw()

```

##### ... by source
```{r, message=FALSE, warning=TRUE}
theta %>% 
  select(- date) %>%
  group_by(source) %>% 
  summarise_all(funs(mean)) -> source_mean

# plot 
melt2 = melt(source_mean, id = "source")

ggplot(melt2, aes(x=source, y = variable)) +   
  geom_raster(aes(fill=value)) + 
  scale_fill_gradient(low = "light blue", high = "dark blue") +
  labs(x="Source", y="Topics", title="Average Topic Weights") + 
  theme_bw()

```

### compute cosine similarity

```{r, message=FALSE}
library(proxy)
cosine_dist <- as.matrix(dist(as.matrix(source_mean[,1:20]), method = "cosine"))
colnames(cosine_dist) <- unique(out$meta$source)
rownames(cosine_dist) <- unique(out$meta$source)
```

### Correlogram
```{r, message=FALSE}
devtools::install_github("sinhrks/ggfortify")
library("ggfortify")
devtools::install_github("kassambara/ggcorrplot")
library(ggcorrplot)
library("ggthemes")

ggcorrplot(cosine_dist, hc.order = TRUE, 
           type = "lower", 
           lab = TRUE, 
           lab_size = 3, 
           method="circle", 
           #colors = c("tomato2", "white", "springgreen3"), 
           title="", 
           ggtheme=theme_bw)
```

### correlations between topics
```{r}
mod.out.corr <- topicCorr(stm)
plot(mod.out.corr)
```

### stmBrowser
```{r}
library(stmBrowser)
setwd(tempdir())
stmBrowser(stm, data=out$meta, c("source","date"), text="text_cleaned")

```