### 2017-10-22 paper/R/webhoseio_explore.Rmd
process webhoseio data

---
title: "Topic Modeling of News"
author: "Franziska Löw"
date: "`r format(Sys.Date())`"
output: github_document
---

```{r, include=FALSE}
# load the packages
libs <- c("tidytext","tidyr","readr","lubridate","tm","stm","RColorBrewer",
          "plyr","dplyr","class","knitr","kableExtra","cldr","data.table",
          "htmlTable","ggplot2","gridExtra","jsonlite","stringr","scales","rjson")
lapply(libs, library, character.only = TRUE)
```

## Load Data
```{r}
rm(list = ls())
load("../output/news_cleaned.Rda")

df %>% filter(site != "zdf.de") -> df
col <- brewer.pal(12, name = "Paired")
```

## Count number of terms in a string
```{r}
df$text_lenght <- sapply(gregexpr("\\S+", df$text), length)
```

## Summary Stats
```{r}
df %>%
  group_by(site) %>%
  summarise(obs = n(),
    fb_shares.mean = mean(fb_shares),
            fb_shares.max = max(fb_shares),
            fb_likes.mean = mean(fb_likes),
            fb_likes.max = max(fb_likes),
            external_links.mean = mean(external_links),
            external_links.max = max(external_links),
            text_length.mean = mean(text_lenght),
            text_length.max = max(text_lenght))
  htmlTable(align = "l")
```

### Text Length
```{r}
ggplot(df, aes(x=as.factor(site), y=text_lenght)) +
  stat_summary(fun.y="median", geom="bar", fill = "blue", alpha = .5) +
  labs(title = "") +
  labs(x="", y="Median Text Length") +
  theme(axis.text.x = element_text(angle = 40, hjust = 1))
```

### Facebook Likes
```{r}
ggplot(df, aes(x=as.factor(site), y=fb_likes)) +
  stat_summary(fun.y="sum", geom="bar", fill = "blue", alpha = .5) +
  labs(title = "") +
  labs(x="", y="Sum of Facebook Likes") +
  theme(axis.text.x = element_text(angle = 40, hjust = 1))
```

```{r}
df$btw <- ifelse(grepl("Bundestagswahl", df$text, ignore.case = TRUE), 1,0)
```

### Number of published Articels by News Source

```{r}
ggplot(df, aes(site)) +
  geom_bar(fill="blue", alpha = .5) +
  xlab("") +
  ylab("Number of Articles") +
  theme(axis.text.x = element_text(angle = 40, hjust = 1))
```

### Share of Articles mentioning Bundestagwahl
```{r}
df %>%
  group_by(site, btw) %>%
  tally() %>%
  group_by(site) %>%
  mutate(n_total = sum(n)) -> count

count %>%
  filter(btw == 1) %>%
  ggplot(aes(site, n/n_total)) +
  geom_col(fill="blue", alpha=.5) +
  xlab("") +
  ylab("Share of Articles mentioning Bundestagwahl") +
  theme(axis.text.x = element_text(angle = 40, hjust = 1))
```

```{r}
btw <- df[which(df$btw == 1),]
```

# Pre-Process Text

```{r, echo=FALSE, message=FALSE, warning=FALSE}
clean.text = function(x)
  {
  x = gsub("DIE WELT", "", x)
  x = gsub("SPIEGEL ONLINE", "", x)
  x = gsub("FOCUS Online", "", x)
  x = gsub("ZEIT ONLINE", "", x)
  x = gsub("stern de", "", x)
  x = gsub("Handelsblatt", "", x)
  x = gsub("Politik Inland Bild de", "", x)
  x = gsub("BILDplus Inhalt", "", x)
  x = gsub("Bild de", "", x)
  x = gsub("SWR Aktuell", "", x)
  x = gsub("SWR de", "", x)
  x = gsub("[ ]NDR[ ]", "", x)
  x = gsub('ZDF \"\"Morgenmagazin\"\"', "", x)
  x = gsub("ZDFmediathek", "", x)
  x = gsub("[ ]ZDF[ ]", "", x)
  x = gsub("[ ]ARD[ ]", "", x)

  x = gsub("die neuesten Nachrichten", "", x)
  x = gsub("RSS Feed", "", x)
  x = gsub("Der Überblick   7 mal 17   Alle Artikel", "", x)
  x = gsub("Der Tag im Überblick", "", x)
  x = gsub("Inhalt Seite 1", "", x)
  x = gsub("Seite 2", "", x)
  x = gsub("dpa", "", x)
  x = gsub("Twitter Anzeige", "", x)
  x = gsub("Anzeige", "", x)
  x = gsub("Foto  \\w+", "", x)
  x = gsub("Video", "", x)
  x = gsub("Quelle \\w+", "", x)
  x = gsub("\\w+ \\w+ Getty Images", "", x)
  x = gsub("AFP", "", x)
  x = gsub("August", "", x)
  x = gsub("Live-Ticker", "", x)
  x = gsub("Deutsche Presse Agentur GmbH   Nachrichtenagentur Alle Beiträge von  anzeigen", "", x)
  x = gsub("Deutsche Presse Agentur", "", x)

  x = gsub("[[:punct:]]", " ", x)  # replace punctuation with space
  x = gsub("[[:cntrl:]]", " ", x)  # replace control characters with space
  x = gsub("[[:digit:]]", "", x)  # remove numbers
  x = gsub("^[[:space:]]+", "", x) # remove whitespace at beginning of documents
  x = gsub("[[:space:]]+$", "", x) # remove whitespace at end of documents 
  x = gsub("[\n{2,}]", "", x)
  x = tolower(x)
  return(x)
}

# apply function to the "orig" dataframe
btw$text_cleaned <- clean.text(btw$text)
```

## Remove Stopwords
```{r}
btw$text_cleaned<- removeWords(btw$text_cleaned, stopwords("german"))
```

## Tokenize terms
```{r}
news_words <- btw %>%
  unnest_tokens(word, text_cleaned) %>%
  count(site, word, sort = TRUE)

## TF-IDF
news_words <- news_words %>%
  bind_tf_idf(word, site, n)

### Terms with high tf-idf
news_words %>%
  #filter(site == "spiegel.de") %>%
  arrange(desc(n))
```

### Add customized stopwords
```{r}
mystopwords <- c("ich","sie","passwort","uhr","bitte","videolänge","dass","mindestens","das","mail","die","zdf","schon","die","neuer abschnitt", "datenschutzerklärung","premium","nutzungsbedingungen","pflichtfelder","registrierung","anzeige","großbuchstaben","sonderzeichen","html","seitennavigation","swr","fullscreen","wochit","statista","club","sagte","mehr")

btw$text_cleaned<- removeWords(btw$text_cleaned, mystopwords)
```

## Tokenize terms
```{r}
news_words <- btw %>%
  unnest_tokens(word, text_cleaned) %>%
  count(site, word, sort = TRUE)

## TF-IDF
news_words <- news_words %>%
  bind_tf_idf(word, site, n)

### Terms with high tf-idf
news_words %>%
  #filter(site == "spiegel.de") %>%
  arrange(desc(n))
```

## Bigrams

```{r}
news_bigrams <- btw %>%
  unnest_tokens(bigram, text_cleaned, token="ngrams", n=2)

news_bigrams %>%
  count(bigram, sort = TRUE)
```

# Topic Modeling
## LDA
```{r}
# Build corpus
btw.lda <- Corpus(VectorSource(btw$text_cleaned))
# Create document-term matrix
dtm <- DocumentTermMatrix(btw.lda)

inspect(dtm)
```

```{r}
# Run LDA using Gibbs sampling
library(topicmodels)

#Set parameters for Gibbs sampling (see https://eight2late.wordpress.com/2015/09/29/a-gentle-introduction-to-topic-modeling-using-r/)
burnin <- 4000
iter <- 2000
thin <- 500
seed <-list(2003,5,63,100001,765)
nstart <- 5
best <- TRUE

k = 10 # Topics
```

```{r}
t1 <- Sys.time()
lda10 <-LDA(dtm, k, method="Gibbs", control=list(nstart=nstart, seed=seed, best=best, burnin=burnin, iter = iter, thin=thin))
t2 <- Sys.time()
t2 - t1
```

```{r}
#write out results
lda10.topics <- as.matrix(topics(lda10))
lda10.topics
```

```{r}
#top 6 terms in each topic
lda10.terms <- as.matrix(terms(lda10,6))
```

```{r}
#probabilities associated with each topic assignment
topicProbabilities <- as.data.frame(lda10@gamma)
```

```{r}
#Find relative importance of top 2 topics
topic1ToTopic2 <- lapply(1:nrow(dtm),function(x)
sort(topicProbabilities[x,])[k]/sort(topicProbabilities[x,])[k-1])
```

```{r}
#Find relative importance of second and third most important topics
topic2ToTopic3 <- lapply(1:nrow(dtm),function(x)
sort(topicProbabilities[x,])[k-1]/sort(topicProbabilities[x,])[k-2])
```

## Structural Topic Model

```{r, eval=FALSE, include=FALSE}
# Process data
processed <- textProcessor(df$text_cleaned, metadata = df[,c("date","site","text_cleaned","text")],
                           wordLengths = c(2,Inf),
                           lowercase = F,
                           removestopwords = F,
                           removenumbers = F,
                           removepunctuation = F,
                           stem = F)
out <- prepDocuments(processed$documents, processed$vocab, processed$meta)
```

### Run the Model 
```{r}

t1 <- Sys.time()
stm <- stm(documents = out$documents, 
           vocab = out$vocab, 
           K=25, 
           prevalence = ~site, 
           content = ~site, 
           data=out$meta, 
           init.type = "Spectral")
t2 <- Sys.time()
t2 - t1

save(stm, out, file = "stm.RDa")
#load(file = "stm.RDa")

```

### Interpret the results

#### Wordclouds by topic
```{r}
cloud(stm, topic =8, scale = c(5,.25))
```

```{r}
# get document - topic distribution 
theta <- as.data.frame(stm$theta)
theta$source <- out$meta$source
theta$date <- out$meta$date
```

#### get the mean distribution of topics  ...
##### ... by date
```{r}
theta %>% 
  select(- source) %>%
  group_by(date) %>% 
  summarise_all(funs(mean)) -> date_mean

# plot 
melt1 = melt(date_mean, id = "date")

ggplot(melt1, aes(x=date, y = variable)) +   
  geom_raster(aes(fill=value)) + 
  scale_fill_gradient(low = "light blue", high = "dark blue") +
  labs(x="Date", y="Topics", title="Average Topic Weights") + 
  theme_bw()

```

##### ... by source
```{r, message=FALSE, warning=TRUE}
theta %>% 
  select(- date) %>%
  group_by(source) %>% 
  summarise_all(funs(mean)) -> source_mean

# plot 
melt2 = melt(source_mean, id = "source")

ggplot(melt2, aes(x=source, y = variable)) +   
  geom_raster(aes(fill=value)) + 
  scale_fill_gradient(low = "light blue", high = "dark blue") +
  labs(x="Source", y="Topics", title="Average Topic Weights") + 
  theme_bw()

```

### compute cosine similarity

```{r, message=FALSE}
library(proxy)
cosine_dist <- as.matrix(dist(as.matrix(source_mean[,1:20]), method = "cosine"))
colnames(cosine_dist) <- unique(out$meta$source)
rownames(cosine_dist) <- unique(out$meta$source)
```

### Correlogram
```{r, message=FALSE}
devtools::install_github("sinhrks/ggfortify")
library("ggfortify")
devtools::install_github("kassambara/ggcorrplot")
library(ggcorrplot)
library("ggthemes")

ggcorrplot(cosine_dist, hc.order = TRUE, 
           type = "lower", 
           lab = TRUE, 
           lab_size = 3, 
           method="circle", 
           #colors = c("tomato2", "white", "springgreen3"), 
           title="", 
           ggtheme=theme_bw)
```

### correlations between topics
```{r}
mod.out.corr <- topicCorr(stm)
plot(mod.out.corr)
```

### stmBrowser
```{r}
library(stmBrowser)
setwd(tempdir())
stmBrowser(stm, data=out$meta, c("source","date"), text="text_cleaned")

```