### 2017-10-25 paper/R/05_lda_eight2late.Rmd
add lda model

---
title: "Topic Modeling of News"
author: "Franziska Löw"
date: "`r format(Sys.Date())`"
output: github_document
---

```{r, include=FALSE}
# load the packages
libs <- c("tidytext","tidyr","readr","lubridate","tm","stm","RColorBrewer","topicmodels",
          "plyr","dplyr","class","knitr","kableExtra","cldr","data.table",
          "htmlTable","ggplot2","gridExtra","jsonlite","stringr","scales","rjson")
lapply(libs, library, character.only = TRUE)
```

## Load Data
```{r}
rm(list = ls())

load("../output/news_cleaned2.Rda")
write.csv(btw, file="../output/btw.csv")

df$date <- as.Date(df$date)

#col <- brewer.pal(12, name = "Paired")
```

## Reduce Dataframe
```{r}
df %>%
  filter(fb_shares > 0) %>%
  filter(!site %in% c("ndr.de","wdr.de")) %>%
  filter(grepl("bundestagswahl", text_cleaned)) -> btw

```

# Topic Modeling
## Build Corpus
```{r}
btw %>%
  filter(nchar(text_cleaned) > 5) -> btw
```

```{r}
# Build corpus
df.lda <- Corpus(VectorSource(btw$text_cleaned))
# Create document-term matrix
dtm <- DocumentTermMatrix(df.lda)
# Assign name to docs
rownames(dtm) <- make.unique(btw$title)

inspect(dtm)
```

## Run LDA using Gibbs sampling
```{r}
#Set parameters for Gibbs sampling (see https://eight2late.wordpress.com/2015/09/29/a-gentle-introduction-to-topic-modeling-using-r/)
burnin <- 1000
iter <- 1000
keep <- 50

k = 50 # Topics
```

```{r}
t1 <- Sys.time()
lda50 <-LDA(dtm, k, method="Gibbs", control=list(burnin=burnin, iter = iter, keep = keep))
t2 <- Sys.time()
t2 - t1

save(ldaOut, file=paste("../output/models/LDAGibbs",k,".Rda"))
```

```{r}
#write out results (docs to topics)
ldaOut.topics <- as.matrix(topics(ldaOut))
write.csv(ldaOut.topics,file=paste("../output/models/LDAGibbs",k,"DocsToTopics.csv"))
```

```{r}
#top 6 terms in each topic
ldaOut.terms <- as.matrix(terms(ldaOut,6))
write.csv(ldaOut.terms,file=paste("../output/models/LDAGibbs",k,"TopicsToTerms.csv"))
```

```{r}
#probabilities associated with each topic assignment
topicProbabilities <- as.data.frame(ldaOut@gamma)
#rownames(topicProbabilities) <- make.unique(btw$title)

write.csv(topicProbabilities,file=paste("../output/models/LDAGibbs",k,"TopicProbabilities.csv"))
```

```{r}
#Find relative importance of top 2 topics
topic1ToTopic2 <- lapply(1:nrow(dtm),function(x)
sort(topicProbabilities[x,])[k]/sort(topicProbabilities[x,])[k-1])
```

```{r}
#Find relative importance of second and third most important topics
topic2ToTopic3 <- lapply(1:nrow(dtm),function(x)
sort(topicProbabilities[x,])[k-1]/sort(topicProbabilities[x,])[k-2])
```

```{r}
#write to file
write.csv(topic1ToTopic2,file=paste("../output/models/LDAGibbs",k,"Topic1ToTopic2.csv"))
write.csv(topic2ToTopic3,file=paste("../output/models/LDAGibbs",k,"Topic2ToTopic3.csv"))
```