### 2017-10-25 paper/R/04_reduced.Rmd

---
title: "Topic Modeling of News"
author: "Franziska Löw"
date: "`r format(Sys.Date())`"
output: github_document
---

```{r, include=FALSE}
# load the packages
libs <- c("tidytext","tidyr","readr","lubridate","tm","stm","RColorBrewer","topicmodels",
          "plyr","dplyr","class","knitr","kableExtra","cldr","data.table",
          "htmlTable","ggplot2","gridExtra","jsonlite","stringr","scales","rjson")
lapply(libs, library, character.only = TRUE)
```

## Load Data
```{r}
rm(list = ls())

load("../output/news_cleaned2.Rda")
write.csv(btw, file="../output/btw.csv")

df$date <- as.Date(df$date)

#col <- brewer.pal(12, name = "Paired")
```

## Reduce Dataframe
```{r}
df %>%
  filter(fb_shares > 0) %>%
  filter(!site %in% c("ndr.de","wdr.de")) %>%
  filter(grepl("bundestagswahl", text_cleaned)) -> btw

```

## Summary Stats
```{r}
btw %>%
  group_by(site) %>%
  summarise(obs = n(),
    fb_shares.mean = mean(fb_shares),
            fb_shares.max = max(fb_shares),
            fb_likes.mean = mean(fb_likes),
            fb_likes.max = max(fb_likes),
            external_links.mean = mean(external_links),
            external_links.max = max(external_links),
            text_length.mean = mean(text_lenght),
            text_length.max = max(text_lenght)) %>%
  htmlTable(align = "l")
```

### Text Length
```{r}
ggplot(btw, aes(x=as.factor(site), y=text_lenght)) +
  stat_summary(fun.y="median", geom="bar", fill = "blue", alpha = .5) +
  labs(title = "") +
  labs(x="", y="Median Text Length") +
  theme(axis.text.x = element_text(angle = 40, hjust = 1))
```

### Facebook Likes
```{r}
ggplot(btw, aes(x=as.factor(site), y=fb_likes)) +
  stat_summary(fun.y="median", geom="bar", fill = "blue", alpha = .5) +
  labs(title = "") +
  labs(x="", y="Median of Facebook Likes") +
  theme(axis.text.x = element_text(angle = 40, hjust = 1))
```

## What are the news with the highest fb-shares?
```{r}
btw %>%
  filter(site == "tagesschau.de") %>%
  arrange(desc(fb_shares)) %>%
  select(text) %>%
  .[1,]
```

```{r}
btw %>%
  ggplot(aes(date, fill = site)) +
  geom_histogram(bins = 100, alpha = .8)
```

## Tokenize
```{r}
btw %>%
  unnest_tokens(word, text_cleaned) %>%
  count(site, word, sort = TRUE) -> news_count

## TF-IDF
news_count <- news_count %>%
  bind_tf_idf(word, site, n)

### Terms with high tf-idf
news_count %>%
  #filter(site == "spiegel.de") %>%
  arrange(desc(tf_idf))
```

### Bigrams
```{r}
btw %>%
  unnest_tokens(bigram, text_cleaned, token="ngrams", n=2) %>%
  count(site, bigram, sort = TRUE) -> news_bigrams

news_bigrams
```

# Topic Modeling
## Build Corpus
```{r}
btw %>%
  filter(nchar(text_cleaned) > 5) -> btw
```

```{r}
# Build corpus
df.lda <- Corpus(VectorSource(btw$text_cleaned))
# Create document-term matrix
dtm <- DocumentTermMatrix(df.lda)

inspect(dtm)
```

## Select Number of K Topics
(http://davidmeza1.github.io/2015/07/20/topic-modeling-in-R.html)

```{r}
seqk <- seq(30, 100, 5) #
burnin <- 1000
iter <- 1000
keep <- 50
system.time(fitted_many <- lapply(seqk, function(k) topicmodels::LDA(dtm, k = k,
                                                     method = "Gibbs",control = list(burnin = burnin,
                                                                         iter = iter, keep = keep) )))
```

### The harmonic mean function
```{r}
harmonicMean <- function(logLikelihoods, precision = 2000L) {
  llMed <- median(logLikelihoods)
  as.double(llMed - log(mean(exp(-mpfr(logLikelihoods,
                                       prec = precision) + llMed))))
}
```

# Run LDA using Gibbs sampling
```{r}
#Set parameters for Gibbs sampling (see https://eight2late.wordpress.com/2015/09/29/a-gentle-introduction-to-topic-modeling-using-r/)
burnin <- 1000
iter <- 1000
keep <- 50

k = 50 # Topics
```

```{r}
t1 <- Sys.time()
lda50 <-LDA(dtm, k, method="Gibbs", control=list(burnin=burnin, iter = iter, keep = keep))
t2 <- Sys.time()
t2 - t1

save(ldaOut, file=paste("../output/models/LDAGibbs",k,".Rda"))
```

```{r}
#write out results (docs to topics)
ldaOut.topics <- as.matrix(topics(ldaOut))
write.csv(ldaOut.topics,file=paste("../output/models/LDAGibbs",k,"DocsToTopics.csv"))
```

```{r}
#top 6 terms in each topic
ldaOut.terms <- as.matrix(terms(ldaOut,6))
write.csv(ldaOut.terms,file=paste("../output/models/LDAGibbs",k,"TopicsToTerms.csv"))
```

```{r}
#probabilities associated with each topic assignment
topicProbabilities <- as.data.frame(ldaOut@gamma)
rownames(topicProbabilities) <- make.unique(btw$title)

write.csv(topicProbabilities,file=paste("../output/models/LDAGibbs",k,"TopicProbabilities.csv"))
```

```{r}
#Find relative importance of top 2 topics
topic1ToTopic2 <- lapply(1:nrow(dtm),function(x)
sort(topicProbabilities[x,])[k]/sort(topicProbabilities[x,])[k-1])
```

```{r}
#Find relative importance of second and third most important topics
topic2ToTopic3 <- lapply(1:nrow(dtm),function(x)
sort(topicProbabilities[x,])[k-1]/sort(topicProbabilities[x,])[k-2])
```

```{r}
#write to file
write.csv(topic1ToTopic2,file=paste("../output/models/LDAGibbs",k,"Topic1ToTopic2.csv"))
write.csv(topic2ToTopic3,file=paste("../output/models/LDAGibbs",k,"Topic2ToTopic3.csv"))
```

## Structural Topic Model

```{r, eval=FALSE, include=FALSE}
# Process data
processed <- textProcessor(df$text_cleaned, metadata = df[,c("date","site","text_cleaned","text")],
                           wordLengths = c(2,Inf),
                           lowercase = F,
                           removestopwords = F,
                           removenumbers = F,
                           removepunctuation = F,
                           stem = F)
out <- prepDocuments(processed$documents, processed$vocab, processed$meta)
```

### Run the Model 
```{r}

t1 <- Sys.time()
stm <- stm(documents = out$documents, 
           vocab = out$vocab, 
           K=25, 
           prevalence = ~site, 
           content = ~site, 
           data=out$meta, 
           init.type = "Spectral")
t2 <- Sys.time()
t2 - t1

save(stm, out, file = "stm.RDa")
#load(file = "stm.RDa")

```

### Interpret the results

#### Wordclouds by topic
```{r}
cloud(stm, topic =8, scale = c(5,.25))
```

```{r}
# get document - topic distribution 
theta <- as.data.frame(stm$theta)
theta$source <- out$meta$source
theta$date <- out$meta$date
```

#### get the mean distribution of topics  ...
##### ... by date
```{r}
theta %>% 
  select(- source) %>%
  group_by(date) %>% 
  summarise_all(funs(mean)) -> date_mean

# plot 
melt1 = melt(date_mean, id = "date")

ggplot(melt1, aes(x=date, y = variable)) +   
  geom_raster(aes(fill=value)) + 
  scale_fill_gradient(low = "light blue", high = "dark blue") +
  labs(x="Date", y="Topics", title="Average Topic Weights") + 
  theme_bw()

```

##### ... by source
```{r, message=FALSE, warning=TRUE}
theta %>% 
  select(- date) %>%
  group_by(source) %>% 
  summarise_all(funs(mean)) -> source_mean

# plot 
melt2 = melt(source_mean, id = "source")

ggplot(melt2, aes(x=source, y = variable)) +   
  geom_raster(aes(fill=value)) + 
  scale_fill_gradient(low = "light blue", high = "dark blue") +
  labs(x="Source", y="Topics", title="Average Topic Weights") + 
  theme_bw()

```

### compute cosine similarity

```{r, message=FALSE}
library(proxy)
cosine_dist <- as.matrix(dist(as.matrix(source_mean[,1:20]), method = "cosine"))
colnames(cosine_dist) <- unique(out$meta$source)
rownames(cosine_dist) <- unique(out$meta$source)
```

### Correlogram
```{r, message=FALSE}
devtools::install_github("sinhrks/ggfortify")
library("ggfortify")
devtools::install_github("kassambara/ggcorrplot")
library(ggcorrplot)
library("ggthemes")

ggcorrplot(cosine_dist, hc.order = TRUE, 
           type = "lower", 
           lab = TRUE, 
           lab_size = 3, 
           method="circle", 
           #colors = c("tomato2", "white", "springgreen3"), 
           title="", 
           ggtheme=theme_bw)
```

### correlations between topics
```{r}
mod.out.corr <- topicCorr(stm)
plot(mod.out.corr)
```

### stmBrowser
```{r}
library(stmBrowser)
setwd(tempdir())
stmBrowser(stm, data=out$meta, c("source","date"), text="text_cleaned")

```.