### 2017-10-31 paper/R/06_explore_LDA.Rmd

---
title: "Topic Modeling of News"
author: "Franziska Löw"
date: "`r format(Sys.Date())`"
output: github_document
---

```{r, include=FALSE}
# load the packages
libs <- c("tidytext","tidyr","readr","lubridate","tm","stm","RColorBrewer","topicmodels",
          "plyr","dplyr","class","knitr","kableExtra","cldr","data.table",
          "htmlTable","ggplot2","gridExtra","jsonlite","stringr","scales","rjson")
lapply(libs, library, character.only = TRUE)
```

## Load Data
```{r}
rm(list=ls())

load(file ="../output/models/LDAGibbs 40 .Rda")
```

### Most likely Topics per Article
```{r}
# The topics function from the package is used to extract the most likely topic for each document 
btw.topics <- topics(ldaOut, 3)

# Create Dataframe
doctopics.df <- as.data.frame(t(btw.topics))

doctopics.df %>%
  transmute(title = rownames(.),
            topic1 = V1,
            Topic2 = V2,
            Topic3 = V3) -> doctopics.df

doctopics.df$ArticleID <- as.integer(rownames(doctopics.df))

# Add Topic to origian DF
btw1 <- btw %>%
  mutate(ArticleID = as.integer(ArticleID)) %>%
  inner_join(.,doctopics.df, by="ArticleID")
```

### Most likely Words per Topic
```{r}
## Return the top 30 terms.
btw.terms <- as.data.frame(terms(ldaOut, 30), stringAsFactors = FALSE)
btw.terms[1:10,10:12]
```

### Create Topic Labels
```{r}
## Return the top 30 terms.
btw.terms <- as.data.frame(terms(ldaOut, 30), stringAsFactors = FALSE)

topicTerms <- btw.terms %>% gather(Topic)
topicTerms <- cbind(topicTerms, Rank = rep(1:30))
topicTerms <- topicTerms %>% filter(Rank < 5)

topicTerms <- topicTerms %>% mutate(topic = stringr::word(Topic, 2))
topicTerms$topic <- as.numeric(topicTerms$topic)

topicLabel <- data.frame()
for (i in 1:40){
  z <- dplyr::filter(topicTerms, topic==i)
  l <- as.data.frame(paste(z[1,2], z[2,2], z[3,2], z[4,2], sep = " "), stringAsFactors = FALSE)
  topicLabel <- rbind(topicLabel, l)
}
colnames(topicLabel) <- c("Label")
topicLabel$topic <- rownames(topicLabel)
```

### Working with Theta (Topic prevalence)
```{r}
theta <- as.data.frame(posterior(ldaOut)$topics)
gamma <- as.data.frame(posterior(ldaOut)$terms)
```

Correlate the topic by news source of the documents. The column means of theta, grouped by news source is calculated.

```{r}
x <- as.data.frame(row.names(theta), stringsAsFactors = FALSE)
colnames(x) <- c("ArticleID")
x$ArticleID <- as.numeric(x$ArticleID)

theta2 <- cbind(x,theta)
theta2 <- dplyr::left_join(theta2, btw[,c("site", "ArticleID")], by="ArticleID")

## Column means grouped by news outlet
theta.mean.by <- by(theta2[,2:51], theta2$site, colMeans)
theta.mean <- do.call("rbind", theta.mean.by)
```

```{r}
library(corrplot)

c <- cor(theta.mean)
corrplot(c, method = "circle")
```

Using the categories mean theta, it is possible to select the most diagnostic topic for each of the categories.

```{r}
theta.mean.ratios <- theta.mean
for (ii in 1:nrow(theta.mean)) {
  for (jj in 1:ncol(theta.mean)) {
    theta.mean.ratios[ii,jj] <-
      theta.mean[ii,jj] / sum(theta.mean[ii,-jj])
  }
}

topics.by.ratio <- apply(theta.mean.ratios, 1, function(x) sort(x, decreasing = TRUE, index.return = TRUE)$ix)

# the most diagnostic topics per news source are found in the thea 1st row of index matrix
topics.by.ratio[1:10,]
```

### compute cosine similarity

```{r, message=FALSE}
library(proxy)
cosine_dist <- as.matrix(dist(topics.by.ratio, by_rows = FALSE, method = "cosine"))
```

### Correlogram
```{r, message=FALSE}
devtools::install_github("sinhrks/ggfortify")
library("ggfortify")
devtools::install_github("kassambara/ggcorrplot")
library(ggcorrplot)
library("ggthemes")

ggcorrplot(cosine_dist, hc.order = TRUE, 
           type = "lower", 
           lab = TRUE, 
           lab_size = 3, 
           method="circle", 
           #colors = c("tomato2", "white", "springgreen3"), 
           title="", 
           ggtheme=theme_bw)
```

## Regression
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
#library(GGally)

btw_reduced <- btw[,c(1,6,10,13,14,15)]

ggpairs(btw_reduced, aes(color = site, alpha=.5))
```

```{r}
library(plm)

lm1 <- lm(fb_shares ~ Topic1 * site, data = btw)
summary(lm1)
```