### 2017-12-08 paper/R/pageImpr.Rmd

---
title: "R Notebook"
output: 
---

```{r}
library(readxl)
library(tidyverse)
```

```{r}
unique_user1 <- read_excel("../data/unique_user1.xls")
unique_user2 <- read_excel("../data/unique_user2.xls")

uu_df <- rbind(unique_user1,unique_user2) %>%
  rename(uniqueUser = `Unique User Tag Mio.`,
         site = Dachmarke,
         date = Tag) %>%
  select(date, site, uniqueUser)
```

```{r}
## Monthly (2003-2017)
unique_user1 <- read_excel("../data/unique_user_2003_2017.xls")
unique_user2 <- read_excel("../data/unique_user_2003_2017_2.xls")

uu_df_full <- rbind(unique_user1,unique_user2) %>%
  mutate(date = as.POSIXct(paste(Jahr,month,"1",sep="-"))) %>%
  rename(uniqueUser = `Unique User Monat Mio.`,
         site = Dachmarke) %>%
  select(date, site, uniqueUser)
```