### 2017-12-08 paper/R/webhose.Rmd

---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
webhose_nov_30 <- read.csv(".../data/eventbride/webhose_nov_30.csv")
```

Test string patterns
```{r}
pattern <- "(?<=main_image': ')[^']*"
str_extract(webhose_nov_30$posts[2],pattern)
```

```{r}
webhose_nov_30$posts[80]
```

```{r}
webhose_nov_30 %>%
  transmute(url = str_extract(posts,"(?<=main_image': ')[^']*"),
         section = str_extract(posts, "(?<=section_title': ')[^']*"),
         site = str_extract(posts,"(?<=site': ')[^']*"),
         title = str_extract(posts, "(?<=title_full': ')[^']*"),
         text = str_extract(posts, "(?<=text': ')[^']*"),
         date = as.POSIXct(str_extract(posts, "(?<=published': ')[^']*"))) %>%
  distinct(title,site, .keep_all = TRUE) %>%
  filter(!is.na(text)) -> df
```

## select politic section
```{r}
pat1 <- "\\w\\.de/politik/deutschland"
pat2 <- "\\w\\.de/politik/inland"
pat3 <- "\\w\\.de/politik/wahl"
pat4 <- "\\w\\.de/politik/bundestagswahl"
pat5 <- "tagesschau\\.de/inland"

df %>%
  filter(grepl(paste(pat1,pat2,pat3,pat4,pat5,
                     sep="|"), url, perl = TRUE)) -> btw
```

```{r}
df %>%
  filter(site == "focus.de") %>%
  select(url) 
```