### 2018-01-19 paper/R/paper.Rmd

---
title: "Regression"
subtitle: ""
author: "Franziska Löw"
date: ""
output: 
  html_document:
    toc: true
    toc_float: true
    theme: "lumen"
    highlight: "tango"
    code_folding: show
    self_contained: true
---

```{r include=FALSE}
library(ggplot2)     # Static data visualization
library(dplyr)       # Data manipulation
library(stringr)     # String manipulation
library(lubridate)   # Date and time manipulation
library(purrr)       # Functional programming
library(tidyr)       # Reshaping
library(magrittr)    # Advanced piping
library(pushoverr)   # Pushover notifications
library(readr)       # Importing data
library(data.table)
library(stm)
library(readxl)

library(igraph)
library(ggpmisc)
library(ggiraph)
library(RColorBrewer) 
library(ggrepel)
#library(plotly)
library(scales)      # Scales
library(viridis)     # Viridis color scales

library(tidytext)    # Tidy text mining
library(stringdist)  # String distances
library(proxy)       # Distance measures

library(htmlwidgets) # JS visuliaztions
library(htmltools)   # Arbitrary html

# Theming
quartzFonts(
  Roboto =
    c("Roboto-Light",
      "Roboto-Bold",
      "Roboto-Regular",
      "Roboto-Thin")
)

theme_set(
  theme_bw(base_family = "Roboto", base_size = 8) +
    theme(
      plot.title = element_text(size = 10,
                                margin = margin(0, 0, 4, 0, "pt")),
      plot.subtitle = element_text(size = 12),
      plot.caption = element_text(size = 6),
      panel.border = element_blank()
    )
)

rm(list=ls())
col <- brewer.pal(5,"Dark2")
```

# The online news market

## Pageviews
```{r}
views1 <- read.csv("../data/alexa/Historical_Traffic_Trends__Pageviews_Percent_1.csv", comment.char="#")
views2 <- read.csv("../data/alexa/Historical_Traffic_Trends__Pageviews_Percent_2.csv", comment.char="#")

views <- left_join(views1, views2 %>% 
                     select(-Metric), by = "Date") %>%
  mutate(date = as.Date(Date, "%m/%d/%Y")) %>%
  select(- c(Metric,Date)) %>%
  gather(key="site", value="views", focus.de:tagesspiegel.de)
```

```{r}
# keeps <- c("spiegel.de", "focus.de", "zeit.de","bild.de",
#           "tagesschau.de")
#drop <- c("stern.de","sueddeutsche.de")

# reach <- reach %>% 
#   mutate(date = as.Date(Date, "%m/%d/%Y")) %>%
#   select(- c(Metric,Date)) %>%
#   gather(key="site", value="reach", focus.de:n.tv.de) %>%
#  filter(!site %in% drop) %>%
#   mutate(site = ifelse(site=="bild.de","Bild.de",site),
#        site = ifelse(site=="focus.de","FOCUS ONLINE",site),
#        site = ifelse(site=="spiegel.de","SPIEGEL ONLINE",site),
#        site = ifelse(site=="zeit.de","ZEIT ONLINE",site),
#        site = ifelse(site=="tagesschau.de","Tagesschau.de",site),
#        site = ifelse(site=="zdf.de","ZDF.de",site),
#        site = ifelse(site=="welt.de","DIE WELT",site),
#        site = ifelse(site=="n-tv.de","n-tv.de",site))

```

```{r}
p1 <- views %>%
  filter(date > as.POSIXct("2017-05-31")) %>%
  ggplot(aes(date, views, color=site)) +
  geom_line(show.legend = F) +
  geom_text_repel(data = views[views$date==max(views$date), ],
                  aes(date, views, label=site),
                  size=2.8, show.legend = F, seed = 5) +
  scale_x_date(limits = c(as.Date("2017-05-31"),as.Date("2018-02-15")),
               breaks = date_breaks("1 month"), 
               labels=date_format("%B %y", tz="CET")) +
  theme(
      #legend.position   = "none",
      legend.text       =  element_text(size=6),
      plot.background   = element_rect("#fafafa", "#fafafa"),
      panel.background  = element_rect("#fafafa"),
      axis.title.x      = element_blank(),
      axis.text.y       = element_text(size = 8)
    ) +
  labs(x="", y="pageviews", color="", title = "", 
      caption = "Source: Alexa Internet, Inc. (www.alexa.com)" )

p1

ggsave(plot= p1,
filename = "../figs/reach.png", device = "png",
width = 6, height = 4,
        dpi = 600
)
```

```{r}
# calculate percentual change
views$perc_change <- (views$views - lag(views$views, n=1)) / (views$views*100)

p1 <- views %>% 
  filter(date >= as.Date("2017-09-17") &
                   date <= as.Date("2017-10-01")) %>%
  ggplot(aes(date, perc_change, color=factor(site))) +
  geom_line(show.legend = F) +
  geom_text_repel(data = views[views$date=="2017-09-24", ],
                  aes(date, perc_change, label=site),
                  size=2.8, show.legend = F, seed = 5) +
  #scale_color_manual(values = brewer.pal(8,"Dark2")) +
  scale_x_date(breaks = date_breaks("1 day"), labels=date_format("%d.%m.", tz="CET")) +
  labs(x="",y="% change", color="") +
  theme(
      legend.text       =  element_text(size=6),
      plot.background   = element_rect("#fafafa", "#fafafa"),
      panel.background  = element_rect("#fafafa"),
      axis.title.x      = element_blank(),
      axis.text.y       = element_text(size = 8)
    )

p1

ggsave(plot= p1,
filename = "../figs/perc_change.png", device = "png",
width = 6, height = 4,
        dpi = 600
)
```

## Traffic source

```{r}
source1 <- read.csv("../data/alexa/Traffic_Sources_1.csv", 
                    comment.char="#", stringsAsFactors = F)
source2 <- read.csv("../data/alexa/Traffic_Sources_2.csv", 
                    comment.char="#", stringsAsFactors = F)

source <- left_join(source1, source2,
                    by=c("Date","Metric")) %>%
  select(- c(Date)) %>%
  gather(key="site", value="traffic", 
         focus.de:tagesspiegel.de)
```

```{r}
p1 <- ggplot(source, aes(site, traffic, fill=Metric,
                   label = round(traffic, digits = 2))) +
  geom_col(alpha = 0.8) +
  geom_text(size = 2, position = position_stack(vjust = 0.5)) +
  labs(x = "", y = "Percentage of total traffic", 
       caption = "Source: Alexa Internet, Inc. (www.alexa.com)") +
  theme(axis.text.x = element_text(angle = 45, size = 7 ))

p1

ggsave(plot= p1,
filename = "../figs/traffic_source.png", device = "png",
width = 6, height = 4,
        dpi = 600
)
```

## Keywords
```{r message=FALSE, warning=FALSE}
setwd("../data/alexa/keywords/")
filenames <- list.files(pattern = ".tsv$")

# create an empty list that will serve as a container to receive the incoming files
list.data<-list()

# create a loop to read in your data
for (i in 1:length(filenames)) {
  list.data[[i]]<-read.delim(filenames[i], comment.char="#")
}

# Create DF
keywords.df <- as.data.frame(list.data[1])

for (i in 2:length(filenames)) {
  keywords.df <- rbind(keywords.df, as.data.frame(list.data[i]))
}
```

```{r}
keepsite <- c("tagesschau.de", "spiegel.de", "zeit.de", "deutschlandfunk.de",
              "zdf.de", "bild.de", "welt.de", "stern.de", "sueddeutsche.de", "focus.de")

keywords.df <- keywords.df %>% 
  filter(nchar(as.character(site))>1)
#  filter(site %in% keepsite)

unique(keywords.df$site)
```

### Search Traffic %
```{r}
p1 <- keywords.df %>%
  arrange(desc(Percent_of_traffic)) %>%
  group_by(site) %>%
  top_n(10, Percent_of_traffic) %>%
  ungroup() %>%
  ggplot(aes(reorder(term, Percent_of_traffic), 
             Percent_of_traffic, fill=site)) +
  geom_col(show.legend = FALSE) +
  theme_bw(base_family = "Roboto", base_size = 4) +
  coord_flip() +
  facet_wrap(~site, scales = "free") +
  labs(x= "", y = "Search traffic %", 
       title = "",
       caption = "Source: Alexa Internet, Inc. (www.alexa.com)")

p1

ggsave(plot= p1,
filename = "../figs/search_traffic.png", device = "png",
width = 7, height = 4,
        dpi = 600
)
```

### Share of voice
```{r}
keyword1 <- c("nachrichten", "news", "bundestagswahl 2017",
            "groko", "jamaika")

keyword2 <- c("cdu","csu","spd","fdp","afd",
           "die linke","die grünen")

```

```{r}
p1 <- keywords.df %>%
  filter(term %in% keyword1) %>%
  ggplot(aes(site,
             Share_of_voice, fill=term)) +
  geom_col(show.legend = F) +
  coord_flip() +
  facet_grid(~term, scales = "free") +
  labs(x= "", y = "Share of voice",
       title = ""
       )

p1

ggsave(plot= p1,
filename = "../figs/keywords1.png", device = "png",
width = 6, height = 4,
        dpi = 600
)
```

```{r}
p1 <- keywords.df %>%
  filter(term %in% keyword2) %>%
  ggplot(aes(site,
             Share_of_voice, fill=term)) +
  geom_col(show.legend = F) +
  coord_flip() +
  facet_grid(~term) +
  labs(x= "", y = "Share of voice",
       caption = "
  Source: Alexa Internet, Inc. (www.alexa.com)
      ",
       title = ""
       )

p1

ggsave(plot= p1,
filename = "../figs/keywords2.png", device = "png",
width = 6, height = 4,
        dpi = 600
)
```

## Audience Network
```{r}
focus <- read.csv("../data/audience_network/Audience_Overlap_1_15_2018_focus.csv", header=FALSE, comment.char="#")
spiegel <- read.csv("../data/audience_network/Audience_Overlap_1_15_2018_spiegel.csv", header=FALSE, comment.char="#")
bild <- read.csv("../data/audience_network/Audience_Overlap_1_15_2018_bildde.csv", header=FALSE, comment.char="#")
zeit <- read.csv("../data/audience_network/Audience_Overlap_1_15_2018_zeitde.csv", header=FALSE, comment.char="#")
tagesschau <- read.csv("../data/audience_network/Audience_Overlap_1_15_2018_tagesschau.csv", header=FALSE, comment.char="#")

zdf <-  read.csv("../data/audience_network/Audience_Overlap_1_15_2018_zdf.csv.crdownload.csv", header=FALSE, comment.char="#")
#stern <- read.csv("../data/audience_network/Audience_Overlap_1_15_2018_stern.csv", header=FALSE, comment.char="#")
welt <- read.csv("../data/audience_network/Audience_Overlap_1_15_2018 _welt.csv", header=FALSE, comment.char="#")
```

```{r}
focus <- focus[-1,1:2] 
colnames(focus) <- c("site","focus.de")

spiegel <- spiegel[-1,1:2] 
colnames(spiegel) <- c("site","spiegel.de")

bild <- bild[-1,1:2] 
colnames(bild) <- c("site","bild.de")

zeit <- zeit[-1,1:2] 
colnames(zeit) <- c("site","zeit.de")

tagesschau <- tagesschau[-1,1:2] 
colnames(tagesschau) <- c("site","tagesschau.de")

zdf <- zdf[-1,1:2] 
colnames(zdf) <- c("site","zdf.de")

welt <- welt[-1,1:2] 
colnames(welt) <- c("site","welt.de")
```

```{r}
# combine dataframes
network_df <- left_join(focus[1:80,], spiegel[1:80,], by="site") %>%
  left_join(., bild[1:80,], by = "site") %>%
  left_join(., zeit[1:80,], by = "site") %>%
  left_join(., tagesschau[1:80,], by = "site") %>%
  left_join(., zdf[1:80,], by = "site") %>%
  left_join(., welt[1:80,], by = "site")

network_df <- na.omit(network_df)

melted <- network_df %>%
    gather(key="seed", value="score", focus.de:welt.de) %>%
  mutate(score = as.integer(score),
         site = as.factor(site),
         seed = as.factor(seed))
```

```{r}
p1 <- ggplot(melted, aes(seed, site, fill=score)) +
  geom_tile() +
  scale_fill_gradient(low="gray95",high=col[3]) +
  scale_x_discrete(labels=c("bilde.de" = "Bild.de", 
                            "focus.de" = "FOCUS ONLINE", 
                            "spiegel.de" = "SPIEGEL ONLINE",
                            "tagesschau.de" = "Tagesschau.de", 
                            "welt.de" ="DIE WELT",
                            "zdf.de" = "ZDF.de",
                            "zeit.de" = "ZEIT ONLINE")) +
  labs(x="", y="",  caption = "
  Source: Alexa Internet, Inc. (www.alexa.com)
      ") +
  theme(
      legend.text       =  element_text(size=6),
      plot.background   = element_rect("#fafafa", "#fafafa"),
      panel.background  = element_rect("#fafafa"),
      axis.title.x      = element_blank(),
      axis.text.x = element_text(size=8, angle = 45),
      axis.text.y       = element_text(size = 8)
    )

ggsave(plot= p1,
filename = "../figs/audience.png", device = "png",
width = 5, height = 6,
        dpi = 600
)

p1
```

# Data

```{r caching, echo = FALSE}
load("../output/btw_combined.Rda")
```

## Distribution of articles
```{r message=FALSE, warning=FALSE}
ggsave({
  btw %>%
  ggplot(aes(site)) +
  geom_bar(fill=col, alpha= .8) +
  labs(x="", y="Number of articles") +
  theme(
      legend.position   = "none",
      plot.background   = element_rect("#fafafa", "#fafafa"),
      plot.title        = element_text(size = 18),
      plot.subtitle     = element_text(size = 8),
      plot.caption      = element_text(size = 7),
      panel.background  = element_rect("#fafafa"),
      axis.title.x      = element_blank(),
      axis.text.y       = element_text(size = 8)
    )

},
filename = "../figs/bar.png", device = "png", 
width = 6, height = 4,
        dpi = 600)
```

```{r message=FALSE, warning=FALSE}
ggsave({
  btw %>%
  group_by(date, site) %>%
  dplyr::summarise(obs = n()) %>%
  ggplot(aes(date, obs, color = site)) +
  geom_line() +
  #geom_vline(aes(xintercept=as.Date("2017-09-24")),
  #           linetype = 2, color="grey40") +
  #geom_vline(aes(xintercept=as.Date("2017-11-20")),
  #               linetype = 2, color="grey40") +
  scale_color_manual(values = col) +
  labs(x="", y="Count",color="") +
  scale_x_date(breaks = date_breaks("1 month"), labels=date_format("%B", tz="CET")) +
  theme(
      #legend.position   = "none",
      plot.background   = element_rect("#fafafa", "#fafafa"),
      plot.title        = element_text(size = 18),
      plot.subtitle     = element_text(size = 8),
      plot.caption      = element_text(size = 7),
      #panel.grid        = element_blank(),
      panel.background  = element_rect("#fafafa"),
      axis.title.x      = element_blank(),
      axis.text.y       = element_text(size = 8)
    )
},
filename = "../figs/timeline.png", device = "png",width = 7, height = 4,
dpi = 600
)

```

# Model Results

```{r caching, echo = FALSE}
#load("../output/models/STM 60 .Rda")
```

```{r Extract wtp and dtp}
# Word-topic probabilities
stmOut %>% tidy("beta") %>% filter(!is.na(topic)) -> posts.wtp
# Document-topic probabilities
stmOut %>% tidy("gamma") -> posts.dtp

##################
# Topic Labeling #
##################
label <- labelTopics(stmOut, n=5)

prob <- as.data.frame(label$prob, stringsAsFactors = F)
frex <- as.data.frame(label$frex, stringsAsFactors = F)
lift <- as.data.frame(label$lift, stringsAsFactors = F)
score <- as.data.frame(label$score, stringsAsFactors = F)

topicLabel <- prob %>%
  transmute(topic = rownames(.),
            topic_name = paste(topic,": ",prob$V1,prob$V2,prob$V3,score$V1,score$V2,score$V3, sep=","),
            prob = paste(prob$V1,prob$V2,prob$V3, sep=","),
            frex = paste(frex$V1,frex$V2,frex$V3, sep=","),
            lift = paste(lift$V1,lift$V2,lift$V3, sep=","),
            score = paste(score$V1,score$V2,score$V3, sep=","))

topicLabel$topic_name <- vapply(lapply(strsplit(topicLabel$topic_name, ","), unique), paste, character(1L), collapse = " ")
rm(prob, frex, lift, score)

####################################################
# Assign topic with highest gamma to each document #
####################################################

top_topics <-
  posts.dtp %>% 
  group_by(document) %>%
  mutate(therank = rank(-gamma, ties.method = "random")) %>%
  filter(therank %in% 1) %>%
  select(- therank)

btw %>%
  mutate(document = articleID) %>%
  inner_join(.,top_topics, by="document") %>%
  ## Combine with Topic label
  left_join(., topicLabel %>%
              select(topic_name, topic) %>%
              mutate(topic = as.numeric(topic)), by="topic") %>%
  filter(gamma >= 0.5) %>%
  mutate(allocation = 1) -> btw
```

After reviewing a sample of articles assigned to the respective topic, we assign a shorter name to the topics based on the most common words in order to improve readability and traceability.

```{r}
btw %>% group_by(topic, site) %>%
  select(topic_name, title) %>% top_n(3) %>%
  arrange(topic) %>%
  htmlTable::htmlTable(align="l")
```

```{r}
topic_name <- c("1: A.Merkel & K.-T.z.Guttenberg","2: SPD (Schulz,Nahles)","3: Election in Niedersachsen","4: Election campaigns", 
                "5: car industry, diesel", "6: exploratory talks CDU/SPD","7: exploratory talks Jamaika",
                "8: people in GER, christmas speech Steinmeier", "9: CSU (Seehofer,Söder)", "10: Federal Election (various)",
                "11: Schulz, TV-Duell", "12: G20 riots", "13: digital & telecommunication topics","14: regional politics", 
                "15: text-processing failure", "16: Turkey (Erdoğan)", "17: Federal statistics", "18: various electoral districts", 
                "19: SPD (Nahles, Geißler, Klingbeil,...)", "20: EU policies, refugee debate", "21: poll results", 
                "22: FDP (C.Lindner)", "23: election campaign A.Merkel (riots finsterwalde)", "24: Armed Forces (v.d.Leyen)", 
                "25: Twitter fail", "26: AfD internal party structures", "27: H.Maas, Facebook, social media", 
                "28: SPD (Schwesig), women, sexism", "29: USA, D.Trump, UN-climate conference", "30: same-sex marriage vote", 
                "31: DIE LINKE", "32: Muslims in GER, anti-Semitism", "33: S.Gabriel, G.Schröder, Russia", 
                "34: AfD (Gauland, Höcke,...)", "35: BAMF, BND, BKA", "36: church, people", "37: political talkshows", 
                "38: legal judgments (linksunten, insider trading, ...)", "39: Bundestag (AfD, Schäuble, Lammert)",
                "40: refugees, deportation", "41: Terrorism in GER (NSA)", "42: refugee cap, family reunion", 
                "43: general statistics", "44: AfD general", "45: RAF, Steinmeier", "46: Glyphosat, Schmidt, (green) policy",
                "47: Terror attacks in GER (Breitscheidplatz)", "48: crime, healthcare statistics", "49: Helmut Kohl", "50: Muslims in GER", "51: SPD, Schulz", "52: Glyphosat scandal",
                "53: Attack in Hamburg", "54: A.Merkel (pension policy)", "55: H.Kohl", "56: Election in Niedersachsen, various", 
                "57: Terrorism, Rock am Ring", "58: Guttenberg, CSU", "59: CDU/SPD coalition", "60: statistics about crimes&terrorism")

topic_name <- as.data.frame(topic_name) %>%
  mutate(topic = seq(1:60))

btw <- btw %>% select(- topic_name) %>%
  left_join(., topic_name, by = "topic")

btw %>% filter(topic==24 & site=="SPIEGEL ONLINE") %>% select(text_cleaned1) %>% .[2,]
```

Figure \ref{fig_topic_proportion} displays the topics ordered by their expected frequency across the corpus. To assign a label to each topic, we used a combination of most frequent words in that topic and words with the highest "score", computed as the log frequency of the word in the topic divided by the log frequency of the word in other topics \citep{roberts_stm:_2016}. The most frequent topic (4) describes the exploratory talks between the CDU/CSU, FDP and Bündnis90/Die Grünen following the federal elections. This is followed by topic 32, which classifies articles that deal with the NSU trial and the main defendant Zschäpe.

```{r}
frequency <- as.data.frame(colMeans(stmOut$theta)) %>% 
  mutate(frequency = colMeans(stmOut$theta),
    topic=as.character(seq(1:40))) %>%
  left_join(., topicLabel %>% select(topic, topic_name),
            by="topic")

ggsave(plot = {
  ggplot(frequency, aes(x=reorder(topic_name, frequency), y=frequency)) + 
    geom_col(fill=col[1], alpha=0.8) +
    coord_flip() +
    labs(x="", y="expected frequency") +
    theme(axis.text = element_text(size=10))
}, filename = "../figs/topic_proportion.png", device = "png",width = 8, height = 12,
dpi = 600)
```

![Topic Proportion](../figs/topic_proportion.png)

## Sample Articles

```{r Document classification}
set.seed(9272)
posts_classification.sdt <-
  posts.dtp %>% 
  #filter(topic %in% keeps) %>%
  group_by(articleID = document) %>% 
  summarise(topic = min(topic[gamma == max(gamma)]), gamma = max(gamma)) %>% 
  ungroup() %>%
  inner_join(btw %>% 
               #filter(topic %in% keeps) %>%
               select(title, title_text, articleID, topic_name, site),
              by = "articleID") %>%
  mutate(topic_title = 
           paste0("Topic ", formatC(topic, flag = "0", width = 2), 
                  " - ", topic_name)) %>%
  group_by(topic_name, site) %>% 
  top_n(10, gamma) %>%
  sample_n(3, replace = TRUE) %>%
  mutate(row = row_number()) %>% 
  ungroup() 
```

```{r}
posts_classification.sdt %>%
  select(site,topic_title,title) %>%
  htmlTable::htmlTable(align="l")
```

## Estimating the Effect of Covariates

### Difference in topic prevalence
The estimateEffect() function explores how prevalence of topics varies across documents according to document covariates (metadata). First, users must specify the variable that they wish to use for calculating an effect. If there are multiple variables specified in estimateEffect(), then all other variables are held at their sample median. These parameters include the expected proportion of a document that belongs to a topic as a function of a covariate, or a first difference type estimate, where topic prevalence for a particular topic is contrasted for two groups.

```{r}
prep <- estimateEffect(1:40 ~site + s(month),
                       stmOut, meta=out_stemmed$meta, uncertainty = "Global")
```

```{r}
summary(prep)
```

Expected difference in topic probability by news source (with 95% Confidence Intervals). 
parameter "covariate": String of the name of the main covariate of interest. Must be enclosed in quotes. All other covariates within the formula specified in estimateEffect will be kept at their median.
```{r}
for (i in 1:40){

  png(paste0("../figs/estimate_effect",i,".png"))
  plot(prep, "site", method = "pointestimate", topics = i,
       labeltype = "custom", custom.labels = unique(prep$data$site),
       ylab = "", xlab = "Mean topic proportion in corpus",
       main = paste0("Topic ",topicLabel$topic_name[i]),
       xlim = c(-.05,0.2))
  dev.off()
}

#plot(prep, "site", method = "pointestimate", topics = keeps[1])
```

```{r}
# for (i in 1:40){
#   plot(prep, "month", method = "continuous", topics = i,
#        printlegend = F,
#        ylab = "", xlab = "Mean topic proportion in corpus",
#        main = paste0("Topic ",i,": ",topicLabel$topic_name[i]))
# }
```

```{r}
p<- posts.dtp %>%
  left_join(., btw %>% select(document, topic_name,
                              site, month),
            by="document")  %>%
  group_by(site, topic, topic_name, month) %>%
  summarize(gamma_mean = mean(gamma)) %>%
  #filter(month!=5) %>%
  ggplot(aes(x=month, y=gamma_mean, fill=factor(topic), text=topic_name)) +
  geom_col() +
  facet_wrap(~site) +
  #facet_grid(~site) +
  theme(legend.position = "none") +
  xlab("Month")

ggplotly(p)
```

What is the distribution of the gamma value?
```{r}
ggsave(plot={
  btw %>%
    #filter(topic %in% keeps) %>%
  ggplot(aes(gamma)) +
  geom_density(fill="blue",alpha=.5,color="blue") 
},
filename = "../figs/gamma_dist.png",
       width = 8, height = 6, device = "png", dpi=600)
```

## Topic Timeline
```{r Data for topic trends chart, message=FALSE, warning=FALSE}
set.seed(7292)
btw %>% 
  mutate(date = as.POSIXct(date),
         allocation = 1) %>%
  mutate(post_period = 
           date %>% with_tz("Europe/Paris") %>% 
           floor_date("week")) %>%
  # Summarise into tidy dataframe
  group_by(post_period, topic, topic_name) %>% dplyr::summarise(articles = sum(allocation)) %>%
  # Mark peaks 
  group_by(topic_name) %>% dplyr::mutate(peak = ifelse(articles==max(articles),1,0)) -> chart_topic_trends.dt
```

```{r message=FALSE, warning=FALSE}
ggsave(
  plot = {
    ggplot(chart_topic_trends.dt, 
       aes(post_period, articles, color=topic_name)) +
  geom_line(show.legend = FALSE) +
  scale_color_manual(
            values = rainbow(60) %>% 
              adjustcolor(red.f = 0.6, green.f = 0.6, blue.f = 0.6)
          ) +
      geom_text_repel(data= subset(chart_topic_trends.dt,peak==1 & articles >=25),
            aes(post_period,articles,label=topic_name), vjust = -1, size = 2,
            size=2.8) +
      geom_vline(aes(xintercept=as.POSIXct("2017-09-23")),
                 linetype = 2, color="grey20") +
      geom_vline(aes(xintercept=as.POSIXct("2017-11-19")),
                 linetype = 2, color="grey20") +
  ylim(c(0,120)) +
    theme(
      legend.position = "none",
      #plot.background   = element_rect("#fafafa", "#fafafa"),
      plot.title        = element_text(size = 18),
      plot.subtitle     = element_text(size = 8),
      plot.caption      = element_text(size = 7),
      panel.grid        = element_blank(),
      #panel.background  = element_rect("#fafafa"),
      axis.title.x      = element_blank(),
      axis.text.y       = element_text(size = 8)
    )
  },
  device = "png",
  filename = "../figs/topic-timeline.png",
  dpi = 600, height = 5, width = 8
)
```

### Topic Distribution by News Page

A common accusation leveled against traditional news media is the amount of bias in reporting various topics. We try to explore the topic distribution of news articles.

```{r}
library(radarchart)
```

```{r}
btw %>% 
  group_by(site, topic, topic_name) %>%
  summarise(articles = sum(allocation)) %>%
  ungroup() %>%
  group_by(site) %>%
  mutate(perc_articles = articles/sum(articles)) %>% 
  ungroup() %>%
  transmute(site=site,
            perc_articles=perc_articles,
            topic=topic_name) %>%
  spread(key=site, value=perc_articles) %>%
  ungroup() %>%
  mutate(topic=factor(topic)) -> radar
```

```{r}
chartJSRadar(scores = radar, labelSize = 9, maxScale = 0.1, 
             showToolTipLabel = TRUE,
             width = 8, height = 8)
```

## Sentiment analysis

```{r}
sent <- c(
  # positive Wörter
  readLines("dict/SentiWS_v1.8c_Negative.txt",
            encoding = "UTF-8"),
  # negative Wörter
  readLines("dict/SentiWS_v1.8c_Positive.txt",
            encoding = "UTF-8")
) %>% lapply(function(x) {
  # Extrahieren der einzelnen Spalten
  res <- strsplit(x, "\t", fixed = TRUE)[[1]]
  return(data.frame(words = res[1], value = res[2],
                    stringsAsFactors = FALSE))
}) %>%
  bind_rows %>% 
  mutate(word = gsub("\\|.*", "", words) %>% tolower,
         value = as.numeric(value)) %>% 
  # manche Wörter kommen doppelt vor, hier nehmen wir den mittleren Wert
  group_by(word) %>% summarise(value = mean(value)) %>% ungroup
```

```{r, message=FALSE, warning=FALSE, include=FALSE}
# Tokenize text
token <- btw %>%
  unnest_tokens(word, text_cleaned1)

# Combine with sentiment values
sentDF <- left_join(token, sent, by="word") %>% 
  mutate(sentiment = as.numeric(value)) %>% 
  filter(!is.na(sentiment)) 
```

```{r fig.height=12}
sentDF %>% group_by(topic_name, site) %>% 
  summarize(sentiment = mean(sentiment)) %>%
  ggplot(aes(topic_name, sentiment, fill=site)) +
  geom_col(fill=col[3], alpha=0.8) +
    facet_wrap(~site, ncol = 1) +
    labs(x="", y="sentiment value") +
    theme(axis.text.x = element_text(size=10, angle = 90))
```

### Radar plot
```{r}
sentDF %>% 
  select(site, topic_name, sentiment) %>%
  spread(key=site, value=sentiment) %>%
  ungroup() -> radar
```

```{r}
chartJSRadar(scores = radar, labelSize = 8, maxScale = 1000, showToolTipLabel = TRUE,
             addDots = FALSE, showLegend = FALSE,
             width = 8, height = 8)
```

### Using sentimentr
```{r}
library(sentimentr)
```

```{r}
# Load dictionaries (from: http://wortschatz.uni-leipzig.de/de/download)
neg_df <- read_tsv("dict/SentiWS_v1.8c_Negative.txt", col_names = FALSE)
pos_df <- read_tsv("dict/SentiWS_v1.8c_Positive.txt", col_names = FALSE)

sentiment_df <- bind_rows(neg_df,pos_df)
names(sentiment_df) <- c("Wort_POS", "polarity", "Inflektionen")

sentiment_df %>% 
  mutate(words = str_sub(Wort_POS, 1, regexpr("\\|", .$Wort_POS)-1),
         words = tolower(words)
         #POS = str_sub(Wort_POS, start = regexpr("\\|", .$Wort_POS)+1)
         ) %>%
  select(words, polarity) -> sentiment_df

sentiment_df <- rbind(sentiment_df, c("nicht",-0.8))

sentiment_df %>% mutate(polarity = as.numeric(polarity)) %>%
  as_key() -> sentiment_df
```

```{r}
btw.reg %>% 
  mutate(text_split = get_sentences(text)) %$%
  sentiment(text_split,
               polarity_dt = sentiment_df) -> btw_sentiment

btw_sentiment_doclevel <- btw_sentiment %>%
  group_by(element_id) %>%
  summarize(sentiment = mean(sentiment))

btw.reg <- left_join(btw.reg, btw_sentiment_doclevel %>%
                       mutate(document = element_id) %>%
                       select(document, sentiment),
                      by="document")
btw.reg %>%
  group_by(site) %>%
  mutate(ave_sentiment = mean(sentiment, na.rm = T)) -> plot_btw
```

```{r}
p1 <- plot_btw %>%
  ggplot(aes(sentiment, factor(site), color=topic_name, 
             text=paste(document,": ",title))) +
  geom_point(alpha=.5, shape=1) +
  xlim(c(-0.25,0.1)) +
  facet_wrap(~topic_name) +
  labs(y="") +
  theme(legend.position = "none")

ggplotly(p1)
```

```{r}
btw.reg %>% 
  filter(document==372) %>%
  mutate(text_split = get_sentences(text)) %$%
  sentiment_by(text_split,
               polarity_dt = sentiment_df) %>%
  sentimentr::highlight()
```

```{r}
btw.reg %>% 
  filter(document==4) %>% select(text_cleaned)
```

### Radar
#### with ggiraphExtra
```{r}
require(ggiraphExtra)
```

```{r}
btw.reg %>%
  group_by(site, topic_name) %>%
  summarize(sentiment = mean(sentiment, na.rm = T)) %>%
  spread(topic_name, sentiment) -> radar
  ungroup() %>%
  mutate(topic=factor(topic_name)) -> radar
```

```{r}
ggsave(plot={
  ggRadar(radar, aes(color=site), rescale = FALSE,
        legend.position = "none") 
  facet_wrap(~topic)
},
file="../figs/radar.png",
dpi=600,
height = 7,
width = 9
)
```

#### with radarchart

```{r}
btw.reg %>% 
  group_by(site, topic, topic_name) %>%
  summarise(sentiment = mean(sentiment, na.rm = TRUE)) %>%
  ungroup() %>%
  transmute(site=site,
            sentiment=sentiment,
            topic=paste(topic, topic_name, sep=": ")) %>%
  spread(key=site, value=sentiment) %>%
  ungroup() %>%
  mutate(topic = factor(topic))-> radar
```

```{r}
chartJSRadar(scores = radar, labelSize = 8, maxScale = 0.1, showToolTipLabel = TRUE,
             addDots = FALSE, showLegend = FALSE,
             width = 8, height = 8)
```