### 2018-01-19 wahlprogramme.Rmd

---
title: "R Notebook"
output: 
---

```{r, include=FALSE}
# load the packages
libs <- c("tidytext","tidyr","readr","lubridate","tm","stm","RColorBrewer",
          "plyr","dplyr","class","knitr","data.table","parallel","SnowballC",
          "htmlTable","ggplot2","gridExtra","jsonlite","stringr","scales","rjson")
lapply(libs, library, character.only = TRUE)
```

# Load and prepare Dataframes
```{r message=FALSE}
rm(list =ls())

load("paper/output/news_cleaned.Rda")
```

# Pre-Process Text
## Artikel über Wahlprogramme
```{r}
pat <- c("Wahlprogramm")

df <- df %>%
  filter(grepl(pat, title))
```

```{r}
df %>% select(title, site) %>% htmlTable::htmlTable(align="l")
```