### 2018-02-02 paper/R/04_Modelresults.Rmd
add l1norm, cosine similarity

---
title: "Regression"
subtitle: ""
author: "Franziska Löw"
date: ""
output: 
  html_document:
    toc: true
    toc_float: true
    theme: "lumen"
    highlight: "tango"
    code_folding: show
    self_contained: true
---

```{r include=FALSE}
library(ggplot2)     # Static data visualization
library(dplyr)       # Data manipulation
library(stringr)     # String manipulation
library(lubridate)   # Date and time manipulation
library(purrr)       # Functional programming
library(tidyr)       # Reshaping
library(magrittr)    # Advanced piping
library(pushoverr)   # Pushover notifications
library(readr)       # Importing data
library(data.table)
library(stm)
library(readxl)
library(Matrix)

library(igraph)
library(corrplot)
library(patchwork)
library(ggpmisc)
library(ggraph)
library(ggiraph)
library(tidygraph)
library(RColorBrewer) 
library(ggrepel)
library(scales)      # Scales
library(viridis)     # Viridis color scales

library(tidytext)    # Tidy text mining
library(stringdist)  # String distances
library(proxy)       # Distance measures

# Theming
quartzFonts(
  Roboto =
    c("Roboto-Light",
      "Roboto-Bold",
      "Roboto-Regular",
      "Roboto-Thin")
)

theme_set(
  theme_bw(base_family = "Roboto", base_size = 8) +
    theme(
      plot.title = element_text(size = 10,
                                margin = margin(0, 0, 4, 0, "pt")),
      plot.subtitle = element_text(size = 12),
      plot.caption = element_text(size = 6),
      plot.background   = element_rect("#fafafa", "#fafafa"),
      panel.background  = element_rect("#fafafa"),
      panel.border = element_blank()
    )
)

rm(list=ls())
col <- brewer.pal(6,"Dark2")
source("func/functions.R")
```

# 2. Data 
```{r caching, echo = FALSE}
#load("../output/btw_combined.Rda")
load("../output/btw_comb_ownership.Rda")
```

## Distribution of articles
```{r message=FALSE, warning=FALSE}
ggsave({
  btw %>%
  ggplot(aes(ownership)) +
  geom_bar(fill=col[1], alpha= .8) +
  labs(x="", y="Number of articles") +
  theme(
      legend.position   = "none"
    )

},
filename = "../figs/bar.png", device = "png", 
width = 6, height = 4,
        dpi = 600)
```

```{r message=FALSE, warning=FALSE}
ggsave({
  btw %>%
  group_by(date) %>%
  dplyr::summarise(obs = n()) %>%
  ggplot(aes(date, obs)) +
  geom_line(color=col[3]) +
  geom_vline(aes(xintercept=as.Date("2017-09-24")),
             linetype = 2, color=col[2]) +
  scale_color_manual(values = col) +
  labs(x="", y="number of articles",color="") +
  scale_x_date(breaks = date_breaks("1 month"), labels=date_format("%B", tz="CET")) +
  theme(
      legend.position   = "none",
      axis.title.x      = element_blank(),
      axis.text       = element_text(size = 8)
    )
},
filename = "../figs/timeline.png", device = "png",width = 7, height = 4,
dpi = 600
)

```

# 3. Model Results
```{r caching, echo = FALSE}
#load("../output/models/STM_stemmed2 40 .Rda")
load("../output/models/STM_ownership 40 .Rda")
out <- out_ownership
```

## Label topics
```{r}
sagelabs <- sageLabels(stmOut, 25)
```

```{r}
stmOut$settings$covariates$yvarlevels
```

### Label topics based on Bild.de

After reviewing a sample of articles assigned to the respective topic, we assign a shorter name to the topics based on the most common words in order to improve readability and traceability.
```{r}
sagelabs$cov.betas[[2]]$problabels[,1:6]
```

```{r}
topics <- matrix(c(1, "Great coalition", 2, "A.Merkel, CSU, Guttenberg", 3, "Jamaica coalition",
                   4, "Diesel scandal", 5, "A.Merkel, glyphosat scandal", 6, "Seehover vs. Söder",
                   7, "political trends after the election", 8, "Merkel vs.Schulz", 9, "Trends in electoral districts", 
                   10, "Steinmeier, Federal President", 11, "G.Schröder & IT themes (Data policy, Wifi, Telekom)", 
                   12, "Federal Election trends", 13, "Climate policy", 14, "G20 Hamburg", 
                   15, "error (text processing failure)", 16, "Federal budget statistics", 
                   17, "Reports on individuals (Berlin)", 18, "Family reunion refugees", 
                   19, "Polls federal election", 20, "AfD, Weidel in social media", 
                   21, "Women empowerment, M.Schwesig (SPD)", 22, "elections in Niedersachsen", 
                   23, "Court sentences, NSU trial", 24, "election capaigns & political talkshows", 
                   25, "German armed forces", 26, "Issues in the Bundestag", 27, "C.Lindner FDP", 
                   28, "Turkey, Russia", 29, "railway problems (DBahn)", 30, "H.Kohl obituary", 
                   31, "AfD Gauland, Höcke", 32, "Asylum claims and deportation", 33, "EU policies", 
                   34, "DIE LINKE", 35, "SPD", 36, "Islam and antisemitism in GER", 37, "AfD in parliament", 
                   38, "Terror attacks (GER)  & police reports", 39, "Church", 40, "RAF"), ncol=2, byrow=T)

topics.df <- as.data.frame(topics) %>%
  transmute(topic_name = paste(V1, V2, sep=": "),
         topic = 1:40) 
```

## Assign a topic to each document (choose by gamma)
```{r Extract wtp and dtp}
# Document-topic probabilities
stmOut %>% tidy("theta") -> theta

top_topics <- theta %>% 
  group_by(document) %>%
  mutate(therank = rank(-gamma, ties.method = "random")) %>%
  filter(therank == 1) %>%
  select(- therank)

btw.2 <- btw %>%
  mutate(document = articleID) %>%
  inner_join(.,top_topics, by="document") %>%
  ## Combine with Topic label
  left_join(., topics.df, by="topic") %>%
  mutate(allocation = 1) 
```

### now we can inspect the unclear topics
```{r}
btw.2 %>% filter(topic==11) %>% select(title, gamma, url) %>%
  htmlTable::htmlTable(align="l")
```

## 3.1. Topic proportions

### in the overall corpus
```{r}
frequency <- as.data.frame(colMeans(stmOut$theta)) %>% 
  mutate(frequency = colMeans(stmOut$theta),
    topic=paste(topics[,1],topics[,2], sep=": "))

p1 <- ggplot(frequency, aes(x=reorder(topic, frequency), y=frequency)) + 
    geom_col(fill=col[1], alpha=0.8) +
    coord_flip() +
    labs(x="", y="expected frequency") +
    theme(axis.text.x = element_text(size=8),
          axis.text.y = element_text(size=11),
          axis.title = element_text(size=10))

ggsave(plot = p1, filename = "../figs/topic_proportion.png", device = "png",width = 8, height = 12,
dpi = 600)
```

### by newspaper
```{r}
# Topic 1
plotNewspaperHistogram(2)
```

### by ownership
plots the expected proportion of topic use in public media minus the expected proportion of topic use in private media. Thus topics more associated with public media appear to the right of zero.

```{r}
freq <- tapply(stmOut$theta[,1], stmOut$settings$covariates$betaindex, mean)
freq <- as.data.frame(freq) %>% 
    mutate(ownership=stmOut$settings$covariates$yvarlevels,
           topic = 1)

for(i in 2:40) {
  freq1 <- tapply(stmOut$theta[,i], stmOut$settings$covariates$betaindex, mean)
  freq1 <- as.data.frame(freq1) %>% 
    transmute(ownership=stmOut$settings$covariates$yvarlevels,
           topic = i,
           freq = freq1)

  freq <- rbind(freq, freq1)
}

freq <- freq %>%
  #mutate(ownership = ifelse(site == "Tagesschau.de", "public", "private")) %>%
  group_by(topic, ownership) %>%
  summarise(freq_mean = mean(freq)) %>%
  spread(ownership, freq_mean) %>%
  mutate(freq_diff = public - private) %>%
  mutate(value = ifelse(freq_diff>=0, 1,0)) %>%
  left_join(., topics.df, by = "topic")
```

```{r}
p1 <- ggplot(freq, aes(reorder(topic_name, freq_diff), freq_diff,
                 fill = factor(value))) +
  geom_col(show.legend = F) +
  scale_fill_manual(values = col[c(1,3)]) +
  coord_flip() +
  theme(axis.text.x = element_text(size=8),
          axis.text.y = element_text(size=11),
          axis.title = element_text(size=10)) +
    labs(x="", y="difference in expected frequency (public-private)")

p1

ggsave(plot = p1, filename = "../figs/topic_proportion_diff.png", device = "png",width = 8, height = 12,
dpi = 600)
```

## 3.2. Difference in topic prevalence
The estimateEffect() function explores how prevalence of topics varies across documents according to document covariates (metadata). First, users must specify the variable that they wish to use for calculating an effect. If there are multiple variables specified in estimateEffect(), then all other variables are held at their sample median. These parameters include the expected proportion of a document that belongs to a topic as a function of a covariate, or a first difference type estimate, where topic prevalence for a particular topic is contrasted for two groups.

```{r}
# effect <- estimateEffect(c(1:40) ~site+s(month), stmOut, 
#                          metadata = out$meta, uncertainty = "None")

effect <- estimateEffect(c(1:40) ~ownership+s(month), stmOut, 
                         metadata = out_ownership$meta, uncertainty = "None")
#save(out_ownership,stmOut, effect, file = "stm_newsdata.RData")
```

```{r}
tables <- vector(mode="list", length = length(effect$topics))

for (i in seq_along(effect$topics)) {
  sims <- lapply(effect$parameters[[i]], function(x) stm:::rmvnorm(500, x$est, x$vcov))
  sims <- do.call(rbind, sims)
  est <- colMeans(sims)
  se <- sqrt(apply(sims,2, stats::var))
  tval <- est/se
  rdf <- nrow(effect$data) - length(est)
  p <- 2*stats::pt(abs(tval), rdf, lower.tail = FALSE)
  topic <- i

  coefficients <- cbind(topic, est, se, tval, p)
  rownames(coefficients) <- attr(effect$parameters[[1]][[1]]$est, "names") 
  colnames(coefficients) <- c("topic", "Estimate", "Std. Error", "t value", "p")
  tables[[i]] <- coefficients
}

out1 <- list(call=effect$call, topics=effect$topics, tables=tables)
```

```{r}
coeff <- as.data.frame(do.call(rbind,out1$tables))

coeff <- coeff %>% 
  mutate(parameter = rownames(coeff),
         parameter = gsub("ownership", "", parameter),
         parameter = ifelse(parameter == "s(month)1", "1_July", parameter),
         parameter = ifelse(parameter == "s(month)2", "2_August", parameter),
         parameter = ifelse(parameter == "s(month)3", "3_September", parameter),
         parameter = ifelse(parameter == "s(month)4", "4_October", parameter),
         parameter = ifelse(parameter == "s(month)5", "5_November", parameter),
         parameter = ifelse(parameter == "s(month)6", "6_December", parameter),
         signifcant = ifelse(p <= 0.5,"yes","no")) %>%
  left_join(., topics.df, by="topic")
```

```{r}
#sites <- c("(Intercept)", "DIE WELT", "FOCUS ONLINE","SPIEGEL ONLINE", "Tagesschau.de","ZEIT ONLINE")
sites <- c("(Intercept)", "public", "private")
p1 <- coeff %>% filter(parameter %in% sites) %>%
  ggplot(aes(x = reorder(topic_name,topic, decreasing=F), y = Estimate, fill=factor(signifcant))) +
  geom_col() +
  scale_fill_manual(values = col[c(2,1)]) +
  coord_flip() +
  facet_wrap(~parameter, ncol = 6) +
  labs(x="", fill="significant at\nthe 5% level") +
  theme(axis.text.x = element_text(angle=90))

p1

# ggsave(plot = p1, filename = "../figs/estimates_site.png", device = "png",width = 9, height = 5,
# dpi = 600)
```

```{r}
p1 <- coeff %>% filter(!parameter %in% sites) %>%
  ggplot(aes(x = reorder(topic_name,topic, decreasing=F), y = Estimate, 
             fill=factor(signifcant))) +
  geom_col() +
  scale_fill_manual(values = col[c(2,1)]) +
  coord_flip() +
  facet_wrap(~parameter, ncol = 6) +
  labs(x="", fill="significant at\nthe 5% level") +
  theme(axis.text.x = element_text(angle=90))

p1

# ggsave(plot = p1, filename = "../figs/estimates_month.png", device = "png",width = 9, height = 5,
# dpi = 600)
```

## 3.3 Topic Correlation
```{r}
## Create Dataframe with correlations
cor <- cor(stmOut$theta[stmOut$settings$covariates$betaindex==1,])
colnames(cor) <- 1:40

df <- as.data.frame(cor) %>% 
  mutate(topic = 1:40) %>%
  melt(., id="topic", value.name="corr") %>%
  mutate(site = stmOut$settings$covariates$yvarlevels[1])

for (i in 2:2) {
  cor <- cor(stmOut$theta[stmOut$settings$covariates$betaindex==i,])
  colnames(cor) <- 1:40

  df1 <- as.data.frame(cor) %>%
    mutate(topic = 1:40) %>%
    melt(., id="topic", value.name="corr") %>%
    mutate(site = stmOut$settings$covariates$yvarlevels[i])

  df <- rbind(df,df1)
}
```

```{r}
df <- df %>% 
  left_join(.,topics.df, by = "topic") %>%
  mutate(topic1 = topic,
         topic_name1 = topic_name,
         topic = as.numeric(variable)) %>%
  select(-topic_name) %>%
  left_join(., topics.df, by="topic")
```

```{r}
p1 <- df %>%
  filter(corr>0.1 & corr < 1) %>%
  ggplot(aes(reorder(topic_name1,topic1), topic_name, 
             fill=corr)) +
  geom_tile() +
  facet_wrap(~site, ncol = 6) +
  #scale_fill_distiller(palette = "Spectral", trans = "reverse") + 
  labs(x="", y="", fill="correlation") +
  theme_bw(base_family = "Roboto", base_size = 10) +
  theme(axis.text.x = element_text(angle=90, size = 6),
        #plot.background = element_rect(fill = 'transparent', colour = 'transparent'),
        strip.background = element_rect(fill = 'transparent', colour = 'transparent'),
        axis.ticks = element_line(colour = 'transparent'))

p1

# ggsave(plot = p1, filename = "../figs/topic_correlation.png", device = "png",width = 9, height = 5,
# dpi = 600)
```

### Network graph
```{r}
i <- 1
cormat <- cor(stmOut$theta[stmOut$settings$covariates$betaindex==i,])
cormat <- ifelse(cormat >=0.05, cor, 0)
# 
g <- igraph::simplify(igraph::graph.adjacency(cormat, mode = "undirected", weight=TRUE))
igraph::V(g)$name <- topics.df$topic_name
igraph::V(g)$props <- colMeans(stmOut$theta)

p1 <- topicCorrNetwork(g, stmOut$settings$covariates$yvarlevels[i])
# 
ggsave(
      paste0("../figs/corrplot",i,".png"),
      p1,
      dpi = 300,
      width = 9,
      height = 6
    )
```

### Overlapping Correlation
Pairs of topics where an edge exists in both Newspaper, we denote with a light blue dot. Pairs of topics where Newspaper1, but not Newspaper2 have an edge between them, we denote with a red square, and those where Newspaper2, but not Newspaper1 have an edge between them we denote with a blue square.
We then sort the matrix by topics that are similarly correlated with other topics in Newspaper1 and Newspaper2 to those that are not similarly correlated.
```{r}
png("../figs/overlapping_correlation.png")
topicCorrCompare(1,5)
dev.off()
```

## 3.4 Distance metric
Compute different metrices and compare them.

Compute the Jensen Shannon divergence of the word-topic probability. The K-by-V (number of words in the vocabulary) matrix logbeta contains the natural log of the probability of seeing each word conditional on the topic.

```{r}
k <- stmOut$settings$dim$K
p <- length(stmOut$settings$covariates$yvarlevels)
```

```{r}
## Find all combinations of ylevel
comb.matrix <- matrix(0, 
                      nrow= p^2, 
                      ncol = 2)

z <- 1
for (i in 1:p) {
  for (j in 1:p) {
    comb.matrix[z,1] <- stmOut$settings$covariates$yvarlevels[i]
    comb.matrix[z,2] <- stmOut$settings$covariates$yvarlevels[j]
    z <- z+1
  }
}
```

### Calculate L1 Norm
```{r}
# Create empty matrix
l1.matrix <- matrix(0, nrow = nrow(comb.matrix), ncol = k)
names(l1.matrix) <- 1:k

# Populate matrix with JSD Values
z <- 1
for (i in 1:p) {

    #print(paste("ylevel = ", i))
    base <- exp(stmOut$beta$logbeta[[i]])
    compare1 <- exp(stmOut$beta$logbeta[[1]])
    compare2 <- exp(stmOut$beta$logbeta[[2]])

  for (j in 1:k) {

    #print(paste("topic = ", j))
    l1.matrix[z,j] <- sum(abs(base[j,]-compare1[j,]))
    l1.matrix[z+1,j] <- sum(abs(base[j,]-compare2[j,]))

  }
    z <- z+p
}

```

```{r}
## Create DF
l1.df <- as.data.frame(l1.matrix) 
colnames(l1.df) <- 1:k

l1.df <- l1.df %>% 
  mutate(News1 = comb.matrix[,1],
         News2 = comb.matrix[,2]) %>%
  melt(.,id=c("News1","News2"), value.name="L1Norm",
       variable.name = "topic") %>%
  mutate(topic = as.integer(topic)) %>%
  left_join(., topics.df, by="topic")
```

### Calculate Cosine Similarity
```{r}
# Create empty matrix
cosine.matrix <- matrix(0, nrow = nrow(comb.matrix), ncol = k)
names(cosine.matrix) <- 1:k

# Populate matrix with JSD Values
z <- 1
for (i in 1:p) {

    #print(paste("ylevel = ", i))
    base <- exp(stmOut$beta$logbeta[[i]])
    compare1 <- exp(stmOut$beta$logbeta[[1]])
    compare2 <- exp(stmOut$beta$logbeta[[2]])

  for (j in 1:k) {

    #print(paste("topic = ", j))
    cosine.matrix[z,j] <- CosSim(base[j,],compare1[j,])
    cosine.matrix[z+1,j] <- CosSim(base[j,],compare2[j,])

  }
    z <- z+p
}

```

```{r}
## Create DF
cosine.df <- as.data.frame(cosine.matrix) 
colnames(cosine.df) <- 1:k

cosine.df <- cosine.df %>% 
  mutate(News1 = comb.matrix[,1],
         News2 = comb.matrix[,2]) %>%
  melt(.,id=c("News1","News2"), value.name="CosineSimilarity",
       variable.name = "topic") %>%
  mutate(topic = as.integer(topic)) %>%
  left_join(., l1.df, by=c("News1","News2","topic"))
```

### Calculate KLD
```{r}
# Create empty matrix
kld.matrix <- matrix(0, nrow = nrow(comb.matrix), ncol = k)
names(kld.matrix) <- 1:k

# Populate matrix with JSD Values
z <- 1
for (i in 1:p) {

    #print(paste("ylevel = ", i))
    base <- exp(stmOut$beta$logbeta[[i]])
    compare1 <- exp(stmOut$beta$logbeta[[1]])
    compare2 <- exp(stmOut$beta$logbeta[[2]])

  for (j in 1:k) {

    #print(paste("topic = ", j))
    kld.matrix[z,j] <- KLD(base[j,],compare1[j,])
    kld.matrix[z+1,j] <- KLD(base[j,],compare2[j,])

  }
    z <- z+p
}
```

```{r}
## Create DF
kld.df <- as.data.frame(kld.matrix) 
colnames(kld.df) <- 1:k

kld.df <- kld.df %>% mutate(News1 = comb.matrix[,1],
         News2 = comb.matrix[,2]) %>%
  melt(.,id=c("News1","News2"), value.name="KLD",
       variable.name = "topic") %>%
  mutate(topic = as.integer(topic)) %>%
  left_join(., cosine.df, by=c("News1","News2","topic"))
```

### Calculate JSD
```{r}
# Create empty matrix
jsd.matrix <- matrix(0, nrow = nrow(comb.matrix), ncol = k)
names(jsd.matrix) <- 1:k

# Populate matrix with JSD Values
z <- 1
for (i in 1:p) {

    #print(paste("ylevel = ", i))
    base <- exp(stmOut$beta$logbeta[[i]])
    compare1 <- exp(stmOut$beta$logbeta[[1]])
    compare2 <- exp(stmOut$beta$logbeta[[2]])

  for (j in 1:k) {

    #print(paste("topic = ", j))
    jsd.matrix[z,j] <- JSD(base[j,],compare1[j,])
    jsd.matrix[z+1,j] <- JSD(base[j,],compare2[j,])

  }
    z <- z+p
}
```

```{r}
## Create DF
jsd.df <- as.data.frame(jsd.matrix) 
colnames(jsd.df) <- 1:k

distance.df <- jsd.df %>% mutate(News1 = comb.matrix[,1],
         News2 = comb.matrix[,2]) %>% 
  melt(.,id=c("News1","News2"), value.name="JSD",
       variable.name = "topic") %>%
  mutate(topic = as.integer(topic)) %>%
  left_join(., kld.df, by=c("News1","News2","topic"))

rm(l1.df, jsd.df, cosine.df, kld.df)
```

### Plot 
```{r}
cols <- rev(brewer.pal(11, 'RdYlBu'))

p1 <- distance.df %>%
  filter(!topic %in% c("11","15")) %>%
  filter(News1!=News2) %>%
  filter(News1=="private") %>%
  ggplot(aes(News1, reorder(factor(topic_name),JSD), fill = JSD)) + 
  geom_tile() +
  scale_fill_gradientn(colours = cols) +
  labs(x="", y="") +
  theme_bw(base_family = "Roboto", base_size = 11) +
  theme(axis.text.x = element_blank(),
        legend.position = "top",
        plot.background = element_rect(fill = 'transparent', colour = 'transparent'),
        strip.background = element_rect(fill = 'transparent', colour = 'transparent'),
        axis.ticks = element_line(colour = 'transparent'))

p2 <- distance.df %>%
  filter(!topic %in% c("11","15")) %>%
  filter(News1!=News2) %>%
  ggplot(aes(News1, reorder(factor(topic_name),JSD), fill = KLD)) + 
  geom_tile() +
  scale_fill_gradientn(colours = cols) +
  labs(x="", y="") +
  theme_bw(base_family = "Roboto", base_size = 11) +
  theme(axis.text.y = element_blank(),
        legend.position = "top",
        plot.background = element_rect(fill = 'transparent', colour = 'transparent'),
        strip.background = element_rect(fill = 'transparent', colour = 'transparent'),
        axis.ticks = element_line(colour = 'transparent'))

p3 <- distance.df %>%
  filter(!topic %in% c("11","15")) %>%
  filter(News1!=News2) %>%
  filter(News1=="private") %>%
  ggplot(aes(News1, reorder(factor(topic_name),JSD), 
             fill = CosineSimilarity)) + 
  geom_tile() +
  scale_fill_gradientn(colours = cols) +
  labs(x="", y="") +
  theme_bw(base_family = "Roboto", base_size = 11) +
  theme(axis.text = element_blank(),
        legend.position = "top",
        plot.background = element_rect(fill = 'transparent', colour = 'transparent'),
        strip.background = element_rect(fill = 'transparent', colour = 'transparent'),
        axis.ticks = element_line(colour = 'transparent'))

p4 <- distance.df %>%
  filter(!topic %in% c("11","15")) %>%
  filter(News1!=News2) %>%
  filter(News1=="private") %>%
  ggplot(aes(News1, reorder(factor(topic_name),JSD), 
             fill = L1Norm)) + 
  geom_tile() +
  scale_fill_gradientn(colours = cols) +
  labs(x="", y="") +
  theme_bw(base_family = "Roboto", base_size = 11) +
  theme(axis.text = element_blank(),
        legend.position = "top",
        plot.background = element_rect(fill = 'transparent', colour = 'transparent'),
        strip.background = element_rect(fill = 'transparent', colour = 'transparent'),
        axis.ticks = element_line(colour = 'transparent'))

(p1 | p2 | p3 | p4 )

ggsave(filename = "../figs/distance.png", device = "png",width = 12, height = 8,
dpi = 600)
```