### 2018-04-05 paper/writing/tables/coeff2.tex
add tables for regression results

% latex table generated in R 3.4.2 by xtable 1.8-2 package
% Thu Apr  5 15:04:12 2018
\begin{table}[ht]
\centering
\resizebox{\textwidth}{!}{
\begin{tabular}{rllrrrr}
  \hline
 & topic\_name & parameter & Estimate & Std. Error & t value & p \\ 
  \hline
70 & 10: M.Schulz, vs. A.Merkel & ZEIT ONLINE & 0.01 & 0.00 & 3.60 & 0.00 \\ 
  71 & 11: politics \& democracy in GER & (Intercept) & 0.01 & 0.00 & 4.88 & 0.00 \\ 
  72 & 11: politics \& democracy in GER & FOCUS ONLINE & 0.00 & 0.00 & 0.92 & 0.36 \\ 
  73 & 11: politics \& democracy in GER & SPIEGEL ONLINE & -0.00 & 0.00 & -1.32 & 0.19 \\ 
  74 & 11: politics \& democracy in GER & stern.de & -0.00 & 0.00 & -1.57 & 0.12 \\ 
  75 & 11: politics \& democracy in GER & Tagesschau.de & -0.00 & 0.00 & -1.79 & 0.07 \\ 
  76 & 11: politics \& democracy in GER & DIE WELT & 0.00 & 0.00 & 1.25 & 0.21 \\ 
  77 & 11: politics \& democracy in GER & ZEIT ONLINE & 0.01 & 0.00 & 3.54 & 0.00 \\ 
  78 & 12: Deportation of criminal Refugees & (Intercept) & 0.03 & 0.00 & 6.49 & 0.00 \\ 
  79 & 12: Deportation of criminal Refugees & FOCUS ONLINE & 0.03 & 0.01 & 4.86 & 0.00 \\ 
  80 & 12: Deportation of criminal Refugees & SPIEGEL ONLINE & 0.00 & 0.01 & 0.19 & 0.85 \\ 
  81 & 12: Deportation of criminal Refugees & stern.de & 0.00 & 0.01 & 0.81 & 0.42 \\ 
  82 & 12: Deportation of criminal Refugees & Tagesschau.de & 0.02 & 0.01 & 3.35 & 0.00 \\ 
  83 & 12: Deportation of criminal Refugees & DIE WELT & 0.02 & 0.01 & 3.21 & 0.00 \\ 
  84 & 12: Deportation of criminal Refugees & ZEIT ONLINE & -0.01 & 0.01 & -0.96 & 0.34 \\ 
  85 & 13: A.Merkel, election campaign & (Intercept) & 0.02 & 0.00 & 9.43 & 0.00 \\ 
  86 & 13: A.Merkel, election campaign & FOCUS ONLINE & 0.00 & 0.00 & 0.55 & 0.58 \\ 
  87 & 13: A.Merkel, election campaign & SPIEGEL ONLINE & 0.00 & 0.00 & 1.01 & 0.31 \\ 
  88 & 13: A.Merkel, election campaign & stern.de & 0.01 & 0.00 & 2.73 & 0.01 \\ 
  89 & 13: A.Merkel, election campaign & Tagesschau.de & -0.01 & 0.00 & -1.71 & 0.09 \\ 
  90 & 13: A.Merkel, election campaign & DIE WELT & 0.00 & 0.00 & 0.33 & 0.74 \\ 
  91 & 13: A.Merkel, election campaign & ZEIT ONLINE & 0.00 & 0.00 & 1.13 & 0.26 \\ 
  92 & 14: Text processing fail & (Intercept) & 0.01 & 0.00 & 9.65 & 0.00 \\ 
  93 & 14: Text processing fail & FOCUS ONLINE & -0.01 & 0.00 & -6.04 & 0.00 \\ 
  94 & 14: Text processing fail & SPIEGEL ONLINE & -0.01 & 0.00 & -7.20 & 0.00 \\ 
  95 & 14: Text processing fail & stern.de & -0.01 & 0.00 & -7.04 & 0.00 \\ 
  96 & 14: Text processing fail & Tagesschau.de & -0.01 & 0.00 & -6.19 & 0.00 \\ 
  97 & 14: Text processing fail & DIE WELT & -0.01 & 0.00 & -6.39 & 0.00 \\ 
  98 & 14: Text processing fail & ZEIT ONLINE & -0.01 & 0.00 & -4.30 & 0.00 \\ 
  99 & 15: Israel, antisemitism, D.Trump & (Intercept) & 0.03 & 0.00 & 8.95 & 0.00 \\ 
  100 & 15: Israel, antisemitism, D.Trump & FOCUS ONLINE & -0.01 & 0.00 & -1.39 & 0.17 \\ 
  101 & 15: Israel, antisemitism, D.Trump & SPIEGEL ONLINE & -0.01 & 0.00 & -1.44 & 0.15 \\ 
  102 & 15: Israel, antisemitism, D.Trump & stern.de & -0.01 & 0.00 & -1.66 & 0.10 \\ 
  103 & 15: Israel, antisemitism, D.Trump & Tagesschau.de & -0.01 & 0.00 & -1.98 & 0.05 \\ 
  104 & 15: Israel, antisemitism, D.Trump & DIE WELT & -0.01 & 0.00 & -1.63 & 0.10 \\ 
  105 & 15: Israel, antisemitism, D.Trump & ZEIT ONLINE & -0.01 & 0.00 & -1.24 & 0.21 \\ 
  106 & 16: Mix: political talkshows, Refugees & (Intercept) & 0.01 & 0.00 & 4.58 & 0.00 \\ 
  107 & 16: Mix: political talkshows, Refugees & FOCUS ONLINE & 0.00 & 0.00 & 1.20 & 0.23 \\ 
  108 & 16: Mix: political talkshows, Refugees & SPIEGEL ONLINE & -0.00 & 0.00 & -1.50 & 0.13 \\ 
  109 & 16: Mix: political talkshows, Refugees & stern.de & -0.00 & 0.00 & -1.31 & 0.19 \\ 
  110 & 16: Mix: political talkshows, Refugees & Tagesschau.de & -0.01 & 0.00 & -2.88 & 0.00 \\ 
  111 & 16: Mix: political talkshows, Refugees & DIE WELT & 0.00 & 0.00 & 2.05 & 0.04 \\ 
  112 & 16: Mix: political talkshows, Refugees & ZEIT ONLINE & -0.00 & 0.00 & -1.54 & 0.12 \\ 
  113 & 17: Debates within SPD & (Intercept) & 0.04 & 0.00 & 10.09 & 0.00 \\ 
  114 & 17: Debates within SPD & FOCUS ONLINE & -0.01 & 0.00 & -1.36 & 0.17 \\ 
  115 & 17: Debates within SPD & SPIEGEL ONLINE & -0.01 & 0.00 & -2.20 & 0.03 \\ 
  116 & 17: Debates within SPD & stern.de & -0.00 & 0.00 & -0.40 & 0.69 \\ 
  117 & 17: Debates within SPD & Tagesschau.de & -0.01 & 0.00 & -3.04 & 0.00 \\ 
  118 & 17: Debates within SPD & DIE WELT & -0.01 & 0.00 & -2.61 & 0.01 \\ 
  119 & 17: Debates within SPD & ZEIT ONLINE & -0.01 & 0.00 & -1.70 & 0.09 \\ 
  120 & 18: Election in Niedersachsen & (Intercept) & 0.02 & 0.00 & 5.93 & 0.00 \\ 
  121 & 18: Election in Niedersachsen & FOCUS ONLINE & -0.00 & 0.00 & -0.72 & 0.47 \\ 
  122 & 18: Election in Niedersachsen & SPIEGEL ONLINE & 0.01 & 0.00 & 1.67 & 0.10 \\ 
  123 & 18: Election in Niedersachsen & stern.de & 0.01 & 0.00 & 1.91 & 0.06 \\ 
  124 & 18: Election in Niedersachsen & Tagesschau.de & 0.00 & 0.00 & 0.46 & 0.65 \\ 
  125 & 18: Election in Niedersachsen & DIE WELT & 0.00 & 0.00 & 0.49 & 0.62 \\ 
  126 & 18: Election in Niedersachsen & ZEIT ONLINE & 0.01 & 0.00 & 2.37 & 0.02 \\ 
  127 & 19: N.Lammert & (Intercept) & 0.00 & 0.00 & 5.17 & 0.00 \\ 
  128 & 19: N.Lammert & FOCUS ONLINE & -0.00 & 0.00 & -1.48 & 0.14 \\ 
  129 & 19: N.Lammert & SPIEGEL ONLINE & -0.00 & 0.00 & -0.71 & 0.48 \\ 
  130 & 19: N.Lammert & stern.de & -0.00 & 0.00 & -1.64 & 0.10 \\ 
  131 & 19: N.Lammert & Tagesschau.de & -0.00 & 0.00 & -2.38 & 0.02 \\ 
  132 & 19: N.Lammert & DIE WELT & -0.00 & 0.00 & -1.80 & 0.07 \\ 
  133 & 19: N.Lammert & ZEIT ONLINE & 0.00 & 0.00 & 1.50 & 0.13 \\ 
  134 & 20: AfD, right-wing radicalism & (Intercept) & 0.02 & 0.00 & 6.12 & 0.00 \\ 
  135 & 20: AfD, right-wing radicalism & FOCUS ONLINE & -0.00 & 0.00 & -0.66 & 0.51 \\ 
  136 & 20: AfD, right-wing radicalism & SPIEGEL ONLINE & -0.00 & 0.00 & -0.75 & 0.45 \\ 
  137 & 20: AfD, right-wing radicalism & stern.de & -0.00 & 0.00 & -0.86 & 0.39 \\ 
  138 & 20: AfD, right-wing radicalism & Tagesschau.de & 0.00 & 0.00 & 0.06 & 0.95 \\ 
  139 & 20: AfD, right-wing radicalism & DIE WELT & 0.00 & 0.00 & 0.49 & 0.62 \\ 
  140 & 20: AfD, right-wing radicalism & ZEIT ONLINE & -0.00 & 0.00 & -0.50 & 0.62 \\ 
   \hline
\end{tabular}
}
\end{table}