### 2018-04-09 paper/writing/tables/coeff2.tex
ccf

% latex table generated in R 3.4.2 by xtable 1.8-2 package
% Mon Apr  9 18:06:44 2018
\begin{table}[ht]
\centering
\resizebox{\textwidth}{!}{
\begin{tabular}{rllrrrr}
  \hline
 & topic\_name & parameter & Estimate & Std. Error & t value & p \\ 
  \hline
57 & 23: CSU, Söder \& Seehover & (Intercept) & 0.04 & 0.00 & 9.36 & 0.00 \\ 
  58 & 23: CSU, Söder \& Seehover & FOCUS ONLINE & -0.00 & 0.00 & -0.74 & 0.46 \\ 
  59 & 23: CSU, Söder \& Seehover & SPIEGEL ONLINE & 0.01 & 0.00 & 1.29 & 0.20 \\ 
  60 & 23: CSU, Söder \& Seehover & stern.de & -0.00 & 0.00 & -0.36 & 0.72 \\ 
  61 & 23: CSU, Söder \& Seehover & Tagesschau.de & -0.01 & 0.01 & -1.13 & 0.26 \\ 
  62 & 23: CSU, Söder \& Seehover & DIE WELT & -0.01 & 0.00 & -1.16 & 0.25 \\ 
  63 & 23: CSU, Söder \& Seehover & ZEIT ONLINE & 0.00 & 0.01 & 0.44 & 0.66 \\ 
  64 & 26: Jamaica fail, Reelections or GroKo? & (Intercept) & 0.03 & 0.00 & 12.15 & 0.00 \\ 
  65 & 26: Jamaica fail, Reelections or GroKo? & FOCUS ONLINE & -0.00 & 0.00 & -1.17 & 0.24 \\ 
  66 & 26: Jamaica fail, Reelections or GroKo? & SPIEGEL ONLINE & -0.00 & 0.00 & -1.12 & 0.26 \\ 
  67 & 26: Jamaica fail, Reelections or GroKo? & stern.de & -0.01 & 0.00 & -3.56 & 0.00 \\ 
  68 & 26: Jamaica fail, Reelections or GroKo? & Tagesschau.de & -0.01 & 0.00 & -2.61 & 0.01 \\ 
  69 & 26: Jamaica fail, Reelections or GroKo? & DIE WELT & -0.01 & 0.00 & -3.37 & 0.00 \\ 
  70 & 26: Jamaica fail, Reelections or GroKo? & ZEIT ONLINE & -0.01 & 0.00 & -1.44 & 0.15 \\ 
  71 & 27: Jamaica Coalition debates & (Intercept) & 0.07 & 0.00 & 14.34 & 0.00 \\ 
  72 & 27: Jamaica Coalition debates & FOCUS ONLINE & -0.03 & 0.01 & -5.20 & 0.00 \\ 
  73 & 27: Jamaica Coalition debates & SPIEGEL ONLINE & -0.01 & 0.01 & -1.40 & 0.16 \\ 
  74 & 27: Jamaica Coalition debates & stern.de & -0.03 & 0.01 & -5.13 & 0.00 \\ 
  75 & 27: Jamaica Coalition debates & Tagesschau.de & -0.02 & 0.01 & -2.75 & 0.01 \\ 
  76 & 27: Jamaica Coalition debates & DIE WELT & -0.01 & 0.01 & -1.38 & 0.17 \\ 
  77 & 27: Jamaica Coalition debates & ZEIT ONLINE & -0.00 & 0.01 & -0.51 & 0.61 \\ 
  78 & 30: AfD, F.Petry \& Meuthen & (Intercept) & 0.03 & 0.00 & 7.87 & 0.00 \\ 
  79 & 30: AfD, F.Petry \& Meuthen & FOCUS ONLINE & -0.00 & 0.00 & -1.25 & 0.21 \\ 
  80 & 30: AfD, F.Petry \& Meuthen & SPIEGEL ONLINE & 0.00 & 0.00 & 0.49 & 0.62 \\ 
  81 & 30: AfD, F.Petry \& Meuthen & stern.de & -0.00 & 0.00 & -0.95 & 0.34 \\ 
  82 & 30: AfD, F.Petry \& Meuthen & Tagesschau.de & -0.01 & 0.00 & -3.23 & 0.00 \\ 
  83 & 30: AfD, F.Petry \& Meuthen & DIE WELT & -0.01 & 0.00 & -2.64 & 0.01 \\ 
  84 & 30: AfD, F.Petry \& Meuthen & ZEIT ONLINE & 0.00 & 0.00 & 0.92 & 0.36 \\ 
  85 & 32: AfD, Gauland \& Weidel & (Intercept) & 0.02 & 0.00 & 6.84 & 0.00 \\ 
  86 & 32: AfD, Gauland \& Weidel & FOCUS ONLINE & 0.00 & 0.00 & 0.14 & 0.89 \\ 
  87 & 32: AfD, Gauland \& Weidel & SPIEGEL ONLINE & -0.00 & 0.00 & -0.17 & 0.86 \\ 
  88 & 32: AfD, Gauland \& Weidel & stern.de & 0.01 & 0.00 & 1.63 & 0.10 \\ 
  89 & 32: AfD, Gauland \& Weidel & Tagesschau.de & -0.01 & 0.00 & -1.37 & 0.17 \\ 
  90 & 32: AfD, Gauland \& Weidel & DIE WELT & 0.01 & 0.00 & 1.49 & 0.14 \\ 
  91 & 32: AfD, Gauland \& Weidel & ZEIT ONLINE & 0.01 & 0.00 & 1.25 & 0.21 \\ 
  92 & 37: AfD \& DIE LINKE in parliament & (Intercept) & 0.02 & 0.00 & 7.01 & 0.00 \\ 
  93 & 37: AfD \& DIE LINKE in parliament & FOCUS ONLINE & 0.01 & 0.00 & 1.16 & 0.25 \\ 
  94 & 37: AfD \& DIE LINKE in parliament & SPIEGEL ONLINE & 0.01 & 0.00 & 2.83 & 0.00 \\ 
  95 & 37: AfD \& DIE LINKE in parliament & stern.de & 0.00 & 0.00 & 0.19 & 0.85 \\ 
  96 & 37: AfD \& DIE LINKE in parliament & Tagesschau.de & 0.01 & 0.00 & 1.09 & 0.28 \\ 
  97 & 37: AfD \& DIE LINKE in parliament & DIE WELT & 0.00 & 0.00 & 0.37 & 0.71 \\ 
  98 & 37: AfD \& DIE LINKE in parliament & ZEIT ONLINE & 0.01 & 0.00 & 2.13 & 0.03 \\ 
  99 & 46: CDU & (Intercept) & 0.02 & 0.00 & 7.35 & 0.00 \\ 
  100 & 46: CDU & FOCUS ONLINE & -0.00 & 0.00 & -0.32 & 0.75 \\ 
  101 & 46: CDU & SPIEGEL ONLINE & 0.00 & 0.00 & 0.71 & 0.48 \\ 
  102 & 46: CDU & stern.de & -0.01 & 0.00 & -1.98 & 0.05 \\ 
  103 & 46: CDU & Tagesschau.de & -0.01 & 0.00 & -3.37 & 0.00 \\ 
  104 & 46: CDU & DIE WELT & -0.00 & 0.00 & -0.52 & 0.60 \\ 
  105 & 46: CDU & ZEIT ONLINE & -0.00 & 0.00 & -0.08 & 0.93 \\ 
   \hline
\end{tabular}
}
\end{table}