### 2018-04-10 paper/writing/tables/coeff2.tex

% latex table generated in R 3.4.2 by xtable 1.8-2 package
% Tue Apr 10 09:26:23 2018
\begin{table}[ht]
\centering
\resizebox{\textwidth}{!}{
\begin{tabular}{rllrrrr}
  \hline
 & topic\_name & parameter & Estimate & Std. Error & t value & p \\ 
  \hline
57 & 23: CSU, Söder \& Seehover & (Intercept) & 0.035 & 0.004 & 9.361 & 0.000 \\ 
  58 & 23: CSU, Söder \& Seehover & FOCUS ONLINE & -0.003 & 0.005 & -0.739 & 0.460 \\ 
  59 & 23: CSU, Söder \& Seehover & SPIEGEL ONLINE & 0.006 & 0.005 & 1.286 & 0.199 \\ 
  60 & 23: CSU, Söder \& Seehover & stern.de & -0.002 & 0.004 & -0.358 & 0.720 \\ 
  61 & 23: CSU, Söder \& Seehover & Tagesschau.de & -0.006 & 0.005 & -1.132 & 0.257 \\ 
  62 & 23: CSU, Söder \& Seehover & DIE WELT & -0.005 & 0.005 & -1.157 & 0.247 \\ 
  63 & 23: CSU, Söder \& Seehover & ZEIT ONLINE & 0.002 & 0.005 & 0.437 & 0.662 \\ 
  64 & 26: Jamaica fail, Reelections or GroKo? & (Intercept) & 0.035 & 0.003 & 12.152 & 0.000 \\ 
  65 & 26: Jamaica fail, Reelections or GroKo? & FOCUS ONLINE & -0.004 & 0.004 & -1.174 & 0.240 \\ 
  66 & 26: Jamaica fail, Reelections or GroKo? & SPIEGEL ONLINE & -0.004 & 0.004 & -1.121 & 0.262 \\ 
  67 & 26: Jamaica fail, Reelections or GroKo? & stern.de & -0.012 & 0.003 & -3.560 & 0.000 \\ 
  68 & 26: Jamaica fail, Reelections or GroKo? & Tagesschau.de & -0.010 & 0.004 & -2.608 & 0.009 \\ 
  69 & 26: Jamaica fail, Reelections or GroKo? & DIE WELT & -0.012 & 0.004 & -3.374 & 0.001 \\ 
  70 & 26: Jamaica fail, Reelections or GroKo? & ZEIT ONLINE & -0.006 & 0.004 & -1.442 & 0.149 \\ 
  71 & 27: Jamaica Coalition debates & (Intercept) & 0.066 & 0.005 & 14.338 & 0.000 \\ 
  72 & 27: Jamaica Coalition debates & FOCUS ONLINE & -0.029 & 0.006 & -5.204 & 0.000 \\ 
  73 & 27: Jamaica Coalition debates & SPIEGEL ONLINE & -0.009 & 0.006 & -1.400 & 0.161 \\ 
  74 & 27: Jamaica Coalition debates & stern.de & -0.028 & 0.005 & -5.130 & 0.000 \\ 
  75 & 27: Jamaica Coalition debates & Tagesschau.de & -0.018 & 0.006 & -2.752 & 0.006 \\ 
  76 & 27: Jamaica Coalition debates & DIE WELT & -0.007 & 0.005 & -1.376 & 0.169 \\ 
  77 & 27: Jamaica Coalition debates & ZEIT ONLINE & -0.003 & 0.006 & -0.509 & 0.611 \\ 
  78 & 30: AfD, F.Petry \& Meuthen & (Intercept) & 0.026 & 0.003 & 7.873 & 0.000 \\ 
  79 & 30: AfD, F.Petry \& Meuthen & FOCUS ONLINE & -0.005 & 0.004 & -1.248 & 0.212 \\ 
  80 & 30: AfD, F.Petry \& Meuthen & SPIEGEL ONLINE & 0.002 & 0.004 & 0.490 & 0.624 \\ 
  81 & 30: AfD, F.Petry \& Meuthen & stern.de & -0.004 & 0.004 & -0.950 & 0.342 \\ 
  82 & 30: AfD, F.Petry \& Meuthen & Tagesschau.de & -0.014 & 0.004 & -3.234 & 0.001 \\ 
  83 & 30: AfD, F.Petry \& Meuthen & DIE WELT & -0.011 & 0.004 & -2.640 & 0.008 \\ 
  84 & 30: AfD, F.Petry \& Meuthen & ZEIT ONLINE & 0.004 & 0.004 & 0.918 & 0.359 \\ 
  85 & 32: AfD, Gauland \& Weidel & (Intercept) & 0.023 & 0.003 & 6.837 & 0.000 \\ 
  86 & 32: AfD, Gauland \& Weidel & FOCUS ONLINE & 0.001 & 0.004 & 0.141 & 0.888 \\ 
  87 & 32: AfD, Gauland \& Weidel & SPIEGEL ONLINE & -0.001 & 0.004 & -0.170 & 0.865 \\ 
  88 & 32: AfD, Gauland \& Weidel & stern.de & 0.007 & 0.004 & 1.627 & 0.104 \\ 
  89 & 32: AfD, Gauland \& Weidel & Tagesschau.de & -0.006 & 0.005 & -1.371 & 0.170 \\ 
  90 & 32: AfD, Gauland \& Weidel & DIE WELT & 0.006 & 0.004 & 1.494 & 0.135 \\ 
  91 & 32: AfD, Gauland \& Weidel & ZEIT ONLINE & 0.006 & 0.005 & 1.246 & 0.213 \\ 
  92 & 37: AfD \& DIE LINKE in parliament & (Intercept) & 0.025 & 0.003 & 7.013 & 0.000 \\ 
  93 & 37: AfD \& DIE LINKE in parliament & FOCUS ONLINE & 0.005 & 0.004 & 1.159 & 0.246 \\ 
  94 & 37: AfD \& DIE LINKE in parliament & SPIEGEL ONLINE & 0.013 & 0.005 & 2.827 & 0.005 \\ 
  95 & 37: AfD \& DIE LINKE in parliament & stern.de & 0.001 & 0.004 & 0.190 & 0.850 \\ 
  96 & 37: AfD \& DIE LINKE in parliament & Tagesschau.de & 0.005 & 0.005 & 1.089 & 0.276 \\ 
  97 & 37: AfD \& DIE LINKE in parliament & DIE WELT & 0.002 & 0.004 & 0.369 & 0.712 \\ 
  98 & 37: AfD \& DIE LINKE in parliament & ZEIT ONLINE & 0.010 & 0.005 & 2.128 & 0.033 \\ 
  99 & 46: CDU & (Intercept) & 0.019 & 0.003 & 7.348 & 0.000 \\ 
  100 & 46: CDU & FOCUS ONLINE & -0.001 & 0.003 & -0.321 & 0.748 \\ 
  101 & 46: CDU & SPIEGEL ONLINE & 0.002 & 0.003 & 0.707 & 0.479 \\ 
  102 & 46: CDU & stern.de & -0.006 & 0.003 & -1.979 & 0.048 \\ 
  103 & 46: CDU & Tagesschau.de & -0.011 & 0.003 & -3.373 & 0.001 \\ 
  104 & 46: CDU & DIE WELT & -0.002 & 0.003 & -0.523 & 0.601 \\ 
  105 & 46: CDU & ZEIT ONLINE & -0.000 & 0.004 & -0.085 & 0.932 \\ 
   \hline
\end{tabular}
}
\end{table}