### 2018-04-15 paper/writing/tables/coeff2.tex

% latex table generated in R 3.4.2 by xtable 1.8-2 package
% Sun Apr 15 20:40:13 2018
\begin{table}[ht]
\centering
\begin{tabular}{rllrrrr}
  \hline
 & topic\_name & parameter & Estimate & Std. Error & t value & p \\ 
  \hline
57 & 23: CSU, Söder \& Seehover & (Intercept) & 0.035 & 0.004 & 8.855 & 0.000 \\ 
  58 & 23: CSU, Söder \& Seehover & FOCUS ONLINE & -0.004 & 0.005 & -0.816 & 0.414 \\ 
  59 & 23: CSU, Söder \& Seehover & SPIEGEL ONLINE & 0.006 & 0.005 & 1.140 & 0.254 \\ 
  60 & 23: CSU, Söder \& Seehover & stern.de & -0.002 & 0.005 & -0.335 & 0.738 \\ 
  61 & 23: CSU, Söder \& Seehover & Tagesschau.de & -0.006 & 0.005 & -1.137 & 0.256 \\ 
  62 & 23: CSU, Söder \& Seehover & DIE WELT & -0.005 & 0.005 & -1.148 & 0.251 \\ 
  63 & 23: CSU, Söder \& Seehover & ZEIT ONLINE & 0.002 & 0.005 & 0.398 & 0.691 \\ 
  64 & 26: Jamaica fail, Reelections or GroKo? & (Intercept) & 0.035 & 0.003 & 12.650 & 0.000 \\ 
  65 & 26: Jamaica fail, Reelections or GroKo? & FOCUS ONLINE & -0.004 & 0.003 & -1.099 & 0.272 \\ 
  66 & 26: Jamaica fail, Reelections or GroKo? & SPIEGEL ONLINE & -0.004 & 0.004 & -1.042 & 0.297 \\ 
  67 & 26: Jamaica fail, Reelections or GroKo? & stern.de & -0.012 & 0.003 & -3.530 & 0.000 \\ 
  68 & 26: Jamaica fail, Reelections or GroKo? & Tagesschau.de & -0.010 & 0.004 & -2.718 & 0.007 \\ 
  69 & 26: Jamaica fail, Reelections or GroKo? & DIE WELT & -0.012 & 0.003 & -3.759 & 0.000 \\ 
  70 & 26: Jamaica fail, Reelections or GroKo? & ZEIT ONLINE & -0.005 & 0.004 & -1.366 & 0.172 \\ 
  71 & 27: Jamaica Coalition debates & (Intercept) & 0.066 & 0.005 & 13.757 & 0.000 \\ 
  72 & 27: Jamaica Coalition debates & FOCUS ONLINE & -0.029 & 0.006 & -5.224 & 0.000 \\ 
  73 & 27: Jamaica Coalition debates & SPIEGEL ONLINE & -0.009 & 0.006 & -1.491 & 0.136 \\ 
  74 & 27: Jamaica Coalition debates & stern.de & -0.028 & 0.006 & -4.955 & 0.000 \\ 
  75 & 27: Jamaica Coalition debates & Tagesschau.de & -0.017 & 0.006 & -2.805 & 0.005 \\ 
  76 & 27: Jamaica Coalition debates & DIE WELT & -0.007 & 0.006 & -1.331 & 0.183 \\ 
  77 & 27: Jamaica Coalition debates & ZEIT ONLINE & -0.003 & 0.007 & -0.529 & 0.597 \\ 
  78 & 30: AfD, F.Petry \& Meuthen & (Intercept) & 0.025 & 0.003 & 7.563 & 0.000 \\ 
  79 & 30: AfD, F.Petry \& Meuthen & FOCUS ONLINE & -0.005 & 0.004 & -1.124 & 0.261 \\ 
  80 & 30: AfD, F.Petry \& Meuthen & SPIEGEL ONLINE & 0.003 & 0.004 & 0.597 & 0.550 \\ 
  81 & 30: AfD, F.Petry \& Meuthen & stern.de & -0.003 & 0.004 & -0.839 & 0.401 \\ 
  82 & 30: AfD, F.Petry \& Meuthen & Tagesschau.de & -0.013 & 0.005 & -2.896 & 0.004 \\ 
  83 & 30: AfD, F.Petry \& Meuthen & DIE WELT & -0.010 & 0.004 & -2.590 & 0.010 \\ 
  84 & 30: AfD, F.Petry \& Meuthen & ZEIT ONLINE & 0.005 & 0.005 & 1.080 & 0.280 \\ 
  85 & 32: AfD, Gauland \& Weidel & (Intercept) & 0.024 & 0.004 & 6.497 & 0.000 \\ 
  86 & 32: AfD, Gauland \& Weidel & FOCUS ONLINE & 0.000 & 0.005 & 0.101 & 0.919 \\ 
  87 & 32: AfD, Gauland \& Weidel & SPIEGEL ONLINE & -0.001 & 0.005 & -0.185 & 0.853 \\ 
  88 & 32: AfD, Gauland \& Weidel & stern.de & 0.007 & 0.004 & 1.538 & 0.124 \\ 
  89 & 32: AfD, Gauland \& Weidel & Tagesschau.de & -0.006 & 0.004 & -1.427 & 0.154 \\ 
  90 & 32: AfD, Gauland \& Weidel & DIE WELT & 0.006 & 0.004 & 1.333 & 0.183 \\ 
  91 & 32: AfD, Gauland \& Weidel & ZEIT ONLINE & 0.006 & 0.005 & 1.125 & 0.261 \\ 
  92 & 37: AfD \& DIE LINKE in parliament & (Intercept) & 0.024 & 0.004 & 6.758 & 0.000 \\ 
  93 & 37: AfD \& DIE LINKE in parliament & FOCUS ONLINE & 0.005 & 0.004 & 1.110 & 0.267 \\ 
  94 & 37: AfD \& DIE LINKE in parliament & SPIEGEL ONLINE & 0.013 & 0.005 & 2.736 & 0.006 \\ 
  95 & 37: AfD \& DIE LINKE in parliament & stern.de & 0.001 & 0.004 & 0.196 & 0.845 \\ 
  96 & 37: AfD \& DIE LINKE in parliament & Tagesschau.de & 0.005 & 0.005 & 1.065 & 0.287 \\ 
  97 & 37: AfD \& DIE LINKE in parliament & DIE WELT & 0.002 & 0.004 & 0.424 & 0.671 \\ 
  98 & 37: AfD \& DIE LINKE in parliament & ZEIT ONLINE & 0.011 & 0.005 & 2.267 & 0.023 \\ 
  99 & 46: CDU & (Intercept) & 0.019 & 0.002 & 8.091 & 0.000 \\ 
  100 & 46: CDU & FOCUS ONLINE & -0.001 & 0.003 & -0.313 & 0.755 \\ 
  101 & 46: CDU & SPIEGEL ONLINE & 0.002 & 0.003 & 0.682 & 0.495 \\ 
  102 & 46: CDU & stern.de & -0.006 & 0.003 & -2.189 & 0.029 \\ 
  103 & 46: CDU & Tagesschau.de & -0.011 & 0.003 & -3.565 & 0.000 \\ 
  104 & 46: CDU & DIE WELT & -0.002 & 0.003 & -0.582 & 0.561 \\ 
  105 & 46: CDU & ZEIT ONLINE & -0.000 & 0.003 & -0.039 & 0.969 \\ 
   \hline
\end{tabular}
\end{table}